---
output:
  bookdown::pdf_document2:
    toc: no
    toc_depth: 4
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
  bookdown::html_document2:
    toc: yes
    toc_depth: 3
  '': default
header-includes:
  - \input{preamble2.tex}
bibliography: references.bib
csl: apa.csl
link-citations: yes
nocite: |
  @worldbank2024_GDP, @worldbank2024, @worldbank2024_Inflation, @worldbank2024_RemittReceived, @focuseconomics2024, @imf2024, @statistaCryptoMarketCap
cache: yes
---

```{=tex}
\pagenumbering{gobble} % Completely disable numbering on title page
\thispagestyle{empty} % Ensure first page has no header/footer
\centering
\LARGE
```
\textbf{Opting Out: Cryptocurrency Under Consideration of Currency Substitution}

\large

Master Thesis

\large

Alec Vayloyan, MSc Student Applied Information and Data Science

\large

Submission for 16.05.2025

\includegraphics[width=3in]{cover.png} \large

Supervisor: Dr. Denis Bieri ([denis.bieri\@hslu.ch](mailto:denis.bieri@hslu.ch){.email})

Co-Supervisor: Dr. Thomas Ankenbrand ([thomas.ankenbrand\@hslu.ch](mailto:thomas.ankenbrand@hslu.ch){.email})

E-Mail Author: [vayloyanalec49\@gmail.com](mailto:vayloyanalec49@gmail.com){.email}

Thesis submitted in partial requirement for a Master's degree in Applied Information and Data Science at the Lucerne University of Applied Sciences

Spring Semester 2025 \vspace{2cm}

```{=tex}
\begin{flushleft}

\includegraphics[width=2.5in]{hslu_logo.png}

\end{flushleft}
\pagenumbering{roman}
\pagestyle{roman}
\newpage
```
\LARGE

```{=tex}
\begin{center}

\Large

\textbf{Abstract}

\end{center}
```
\normalsize

We analyze the drivers behind this shift, including trust in traditional banking, regulatory landscapes, and the increasing adoption of blockchain technology. Using a combination of quantitative data analysis and qualitative insights, we examine whether Bitcoin and other cryptocurrencies serve as viable alternatives to government-issued money. The study also investigates the risks associated with cryptocurrency-based economies, including price volatility, regulatory uncertainty, and financial exclusion, this is a fake abstract placeholder.

\vspace{8cm}

\raggedright

\textbf{Keywords:} Cryptocurrency, Currency Substitution, Dollarization 2.0, Bitcoin

\newpage

\tableofcontents

\newpage

\listoffigures

\clearpage

\listoftables

\newpage

```{=tex}
\pagenumbering{arabic}
\pagestyle{arabic}
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r getting libraries, include = F, warning = F}
library(dplyr)
library(rvest)
library(xml2)
library(httr)     # OpenAI Access
library(jsonlite) # OpenAI Access
library(mice)     # Evaluating Missings
library(tidyr)
library(tibble)
library(openxlsx)
library(kableExtra)
library(knitr)
library(bookdown)
library(openxlsx)
library(ggplot2)
library(lubridate)
library(countrycode)
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggridges)
library(reshape2)
library(gt)
library(broom)
library(caret) #YJ Transformation
```

```{r setting working directory, include = F,}
setwd("C:/Users/vaylo/OneDrive/Desktop/Masters Thesis/Working Directory/Data")
```

```{r d.Country_Codes, include = F }
# Loading World Bank country codes from a website
url <- "https://irows.ucr.edu/research/tsmstudy/wbcountrycodes.htm" 
html_content <- read_html(url)

d.Country_Codes <- html_content %>%
  html_node("table") %>%  
  html_table()

#Cleaning Up
d.Country_Codes <- d.Country_Codes %>% 
  rename(Country.Name = X1, Country.Code = X2) %>% 
  slice(-(1:2)) %>% 
  filter(!row_number() %in% c(4, 9, 21, 62, 66, 72, 74, 76, 77, 79, 118,     131, 132))


v.Country_Codes <- d.Country_Codes$Country.Code
remove(url,html_content)

# New country codes to add - some were missing
new_country_codes <- c("SRB", "ROU", "SGP", "ARE")  # Add as many as needed

# Append new codes to the existing vector
v.Country_Codes <- unique(c(v.Country_Codes, new_country_codes))  # Ensure uniqueness
```

```{r indicator list, include = F}
d.Indicators <- data.frame(
  Indicator = character(),
  Indicator_Code = character()
)

unique_combinations <- function(df, col1_name, col2_name) {
  # Get the unique values of both specified columns
  col1_values <- unique(df[[col1_name]])
  col2_values <- unique(df[[col2_name]])
  
  # Create a data frame with all unique combinations of the two columns
  combinations_df <- expand.grid(Col1 = col1_values, Col2 = col2_values)
  
  return(combinations_df)
}
```

```{r d.Inflation, include = F}
# Working with Inflation -> 
# Source: https://data.worldbank.org/indicator/FP.CPI.TOTL.ZG

path <- "Data/d.Inflation/API_FP.CPI.TOTL.ZG_DS2_en_csv_v2_3401965.csv"
d.Inflation <- read.csv(path, header = T, skip = 4) 

# Extracting indicators and putting them in a separate running tally 
indicators <- unique_combinations(d.Inflation, "Indicator.Name", "Indicator.Code")
colnames(indicators) <- colnames(d.Indicators)
d.Indicators <- rbind(d.Indicators, indicators)
remove(indicators)

# Cleaning Data d.Inflation
d.Inflation <- d.Inflation %>%
  filter(Country.Code %in% v.Country_Codes) %>%
  select(Country.Code, X2019:X2023)  # Removed Indicator.Code

# Rename first column to "Country"
colnames(d.Inflation)[1] <- "Country"

# Remove 'X' prefix from year columns
colnames(d.Inflation) <- gsub("^X", "", colnames(d.Inflation))

d.Inflation <- d.Inflation # %>%
  #mutate(`2024` = NA)
```

```{r, d.GDP, include = F}
# Working with GDP per Capita -> 
# Source: https://data.worldbank.org/indicator/NY.GDP.PCAP.CD

path <- "Data/d.GDP/API_NY.GDP.PCAP.CD_DS2_en_csv_v2_3401556.csv"
d.GDP <- read.csv(path, header = T, skip = 4) 

# Extracting indicators and putting them in a separate running tally 
indicators <- unique_combinations(d.GDP, "Indicator.Name", "Indicator.Code")
colnames(indicators) <- colnames(d.Indicators)
d.Indicators <- rbind(d.Indicators, indicators)
remove(indicators)

# Cleaning Data d.GDP
d.GDP <- d.GDP %>%
  filter(Country.Code %in% v.Country_Codes) %>%
  select(Country.Code, X2019:X2023)  # Removed Indicator.Code

# Rename first column to "Country"
colnames(d.GDP)[1] <- "Country"

# Remove 'X' prefix from year columns
colnames(d.GDP) <- gsub("^X", "", colnames(d.GDP))

d.GDP <- d.GDP #%>%
  #mutate(`2024` = NA)
```

```{r d.GDS, include = F}
# Working with Gross Domestic Savings (% of GDP) -> 
# Source: https://data.worldbank.org/indicator/NY.GDS.TOTL.ZS

path <- "Data/d.Savings/API_NY.GDS.TOTL.ZS_DS2_en_csv_v2_1988.csv"
d.GDS <- read.csv(path, header = T, skip = 4) 

# Extracting indicators and putting them in a separate running tally 
indicators <- unique_combinations(d.GDS, "Indicator.Name", "Indicator.Code")
colnames(indicators) <- colnames(d.Indicators)
d.Indicators <- rbind(d.Indicators, indicators)
remove(indicators)

# Cleaning Data d.GDS
d.GDS <- d.GDS %>%
  filter(Country.Code %in% v.Country_Codes) %>%
  select(Country.Code, X2019:X2023)  # Removed Indicator.Code

# Rename first column to "Country"
colnames(d.GDS)[1] <- "Country"

# Remove 'X' prefix from year columns
colnames(d.GDS) <- gsub("^X", "", colnames(d.GDS))

#d.GDS <- d.GDS #%>%
  #mutate(`2024` = NA)
```

```{r d.RR, include = F}
# Working with Personal Remittances Received (% of GDP) -> 
# Source: https://data.worldbank.org/indicator/BX.TRF.PWKR.DT.GD.ZS

path <- "Data/d.RemittancesGDP/API_BX.TRF.PWKR.DT.GD.ZS_DS2_en_csv_v2_3426.csv"
d.RR <- read.csv(path, header = T, skip = 4) 

# Extracting indicators and putting them in a separate running tally 
indicators <- unique_combinations(d.RR, "Indicator.Name", "Indicator.Code")
colnames(indicators) <- colnames(d.Indicators)
d.Indicators <- rbind(d.Indicators, indicators)
remove(indicators)

# Cleaning Data d.RR
d.RR <- d.RR %>%
  filter(Country.Code %in% v.Country_Codes) %>%
  select(Country.Code, X2019:X2023)  # Removed Indicator.Code

# Rename first column to "Country"
colnames(d.RR)[1] <- "Country"

# Remove 'X' prefix from year columns
colnames(d.RR) <- gsub("^X", "", colnames(d.RR))

d.RR <- d.RR #%>%
  #mutate(`2024` = NA)
```

```{r, d.External_Debt, include = F, warning = F}
# Working with External Debt (% of GDP) -> 
# Source: https://www.focus-economics.com/economic-indicator/external-debt/ 

path <- "Data/d.External_Debt_GDP/d.External_Debt_GDP.xlsx"

# Loading Raw Data
d.ED <- read.xlsx(path)

# Adding World Bank Codes

d.ED$Country <- countrycode(d.ED$Country, "country.name", "iso3c")
# warning about unconnected code can be safely ignored since the dataset d.Adoption does not contain the two unmatched countries (Ivory Coast, Kosovo) anyways.

#for some reason, column 2,3 did not save as numeric
d.ED$'2019' <- as.numeric(d.ED$'2019')
d.ED$'2020' <- as.numeric(d.ED$'2020')

d.ED <- d.ED #%>%
  #mutate(`2024` = NA)
```

```{r AdoptionLoad, echo = F}
path <- "Data/d.Adoption/Statista2024.xlsx"
d.Adoption <- read.xlsx(path, sheet = "Data")

# Convert country names to ISO3 codes
d.Adoption$Country <- countrycode(d.Adoption$Country, "country.name", "iso3c")
```

# Motivation and Topic Definition

The development of cryptocurrencies, spawned alongside blockchain technology, have challenged traditional monetary systems and economic structures. Since its inception in the wake of the Global Financial Crisis, cryptocurrencies and particularily Bitcoin have been championed by advocates for its decentralized nature, security, and potential as a replacement for fiat currencies. Cryptocurrencies are often viewed as a solution for individuals and nations where conventional financial systems are dysfunctional or highly volatile, providing an alternative currency that is not controlled by the same rules as a potentially dysfunctional monetary system.

A definition of cryptocurrencies, is required. @arslanian2022 provides a useful taxonomy of digital assets. With cryptocurrencies, this paper means an digital asset falling into the bracket of being "**fungible**" and in the sub-bracket "**payment** **token**". Fungibility refers to individual tokens being functionally equivalent - there is no reason to prefer one token of a certain digital asset over another. This is similar to how no one pays attention about the serial number of a fiat bill. Within the fungible category, there are 3 categories, of which only the payment token is relevant for this paper. These are those fungible crypto assets designed to fulfill the functions of money (medium of exchange, store of value, unit of account). These tokens share the following general characteristics: time and location independent availability, (theoretical) security, speed, low-fee transactions, irreversible transactions, (pseudo) anonymity. Three distinct sub-categories of these tokens are now discussed.

## "Traditional" Cryptocurrencies

In contrast to fiat currencies, which derive their value from governmental decree and are backed by legal frameworks, traditional cryptocurrencies operate on a decentralized network of peer-to-peer transactions that do not rely on a central authority. This decentralized nature means that transactions are verified by a distributed ledger known as blockchain, rather than by a trusted intermediary like a bank. Some cryptocurrencies, most notably Bitcoin have a function which reduces and eventually ceases the issuance of new coins through network design. This theoretically deflationary approach is in stark contrast to fiat currencies, where the authorities usually target an inflation rate of around 2% annually [@ammous2018; @centralbanknews2025]. The slow(ing) growth in supply of Bitcoin relative to the existing stock of coins is what leads to idea that Bitcoin is "Digital Gold", since the stability in value of gold is attributable to the same phenomenon of large existing stocks with very slow growth in supply [@ammousBTCstd]. The lack of control by a single entity and slow growth have spawned the belief that cryptocurrencies are immune to inflationary pressures, government control, and political instability, which can all negatively affect those using fiat currencies in the *status quo* financial system.

## Stablecoins

An important subset of payment tokens in relation to this topic are stablecoins. These are cryptocurrencies which have prices linked to reference assets. In this sense, they are different to "traditional" cryptocurrencies in that the price is driven by centralized actors, rather than the collective decisions of the users. @catalini2022 identify the main ways stability can be achieved are:

1.  Backing of the stablecoins by one or more fiat currencies (most common).

2.  Backing of the stablecoin by one or more cryptocurrencies, not issued by the same entity as the stablecoin.

3.  Backing of the stablecoin by one or more cryptocurrency issued by the same entity as the stablecoin, which can be used to manipulate the price of the stablecoin (algorithmic stablecoins).

4.  A suggested market mechanism for stability has also been found in the research. Some papers argue that the increased purchase of stablecoins during market downturns, rather than traditional cryptocurrencies acts as an additional stabilizing force for stablecoin's prices, beyond the currency architecture [@baur2021; @lyons2020].

Stablecoins have the potential to combine the stability benefits of well managed fiat currencies with particularily the settlement speed advantages of payment tokens [@catalini2022]. There is evidence that these tokens are already being used precisely for this. For example, a local Brazilian crypto expert spoke to Chainalysis on the large transaction volumes on Brazilian exchanges, stating "many of Brazil's exchanges and fintech brokerages offer USD-pegged stablecoins to their customers\...but at this stage, it appears that the main use cases for stablecoins are on the B2B[^1] cross-border payments side" [Aaron Stanley as cited in @chainalysis2024 , p. 33]. While the consensus is that stablecoins are more stable than traditional cryptocurrencies, they have not managed to maintain the desired pegs to fiat currencies consistently [@baughman2022; @kosse2023]. Despite these issues, the interest in stablecoins has increased from June 2023 - June 2024 (albeit less strongly compared to traditional cryptocurrencies) among institutions making transactions under USD 10M, with the share of stablecoin inflows on exchanges outside the US increasing [@chainalysis2024]. All of this is an indication that the stablecoin subset of cryptocurrencies will continue be relevant in the future as both an investment vehicle and a means to cross borders.

[^1]: B2B: Business-to-Business (not in original quote)

## Central Bank Digital Currencies

Central Bank Digital Currencies are "a new form of digitized sovereign currency, generally conceived to be equal or physical cash or reserves held at the central bank" [@arslanian2022, p. 171]. While the interest in such technologies is increasing drastically, with most central banks (including every G20 country's) exploring the issue, no major economy has created and fully launched one - at the of writing just the Bahamas, Jamaica and Nigeria have done so, with some other countries having advanced pilot programs which already had transactions [@atlanticcouncil2025]. Central Bank Digital Currencies have many theoretical advantages as well as risks [@arslanian2022; @genc2024]. An extensive discussion of these is beyond the scope of this paper, however it is the same government involvement in a digital asset which can bring substantial benefit also comes with a level of control by a small group of policymakers that brings in itself the risk of deliberate misuse against public interest or mismanagement.

## Adoption of Cryptocurrencies

Cryptocurrency adoption has not been uniform. As Figure \@ref(fig:AdoptionMap) shows, for those countries with available data, the number of survey respondents who said they use cryptocurrency was as low as 6% and as high as 47%. That is a considerable difference! Factors driving cryptocurrency adoption vary greatly across countries and are heavily debated among academics (see \@ref(adoption-of-cryptocurrency)). Understanding the global drivers of cryptocurrency adoption, particularly in the context of factors that are more relevant for emerging markets with less developed financial systems provides a valuable insight into cryptocurrency's potential future size and role in the global economy, particularily as the last couple of years have seen substantial growth in the interest coming from countries outside the high - income bracket [@chainalysis2024].

\FloatBarrier

```{r AdoptionMap, echo = F, fig.cap="Worldmap Showing Cryptocurrency Adoption Percentage for 2023 (Statista, 2024a)",fig.width=6, fig.height=4, fig.pos = 'H'}
selected_year <- "2023"  # Changeable variable
# Rename columns for compatibility
d.Adoption_long <- d.Adoption %>%
  rename(iso_a3 = Country) %>%
  select(iso_a3, all_of(selected_year))

# Load world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# Merge dataset with world map
world_data <- left_join(world, d.Adoption_long, by = "iso_a3")

# Plot world map
ggplot(data = world_data) +
  geom_sf(aes(fill = get(selected_year))) +
  scale_fill_gradientn(colors = c("purple", "green", "yellow"), na.value = "gray") +
  labs(#title = paste("Adoption Data for Year", selected_year),
       fill = "% Adoption") +
  theme_minimal()
```

\FloatBarrier

## Decentralized Alternatives

Two key factors have led to the idea that cryptocurrencies can be used as an alternative currency by people in dysfunctional financial systems. These two factors are the **internet-based nature of cryptocurrencies** and the **relative price stability of cryptocurrencies.**

Firstly, the internet based nature means that most people with a smartphone and internet can relatively easily access the necessary infrastructure to buy and sell cryptocurrencies. This extends beyond political borders - there is nothing intrinsically hindering people or institutions in different political jurisdictions from exchanging with each other. This is different to using fiat currencies, where financial institutions (usually) comply with regulations on the transfer of digital funds and the lack of physical proximity between buyers and sellers fiat currency limits the ability to exchange cash.

Secondly, depending on the inflationary context, the price of cryptocurrencies can be relatively stable. Stability of a local currency is usually measured either against inflation or a exchange rate to a major currency. The price fluctuations of particularily stablecoins can be lower than those of many fiat currencies, and even Bitcoin, which has no intrinsic stability mechanisms beyond the deflationary architecture, has maintained it's value against fiat currencies in high inflation settings.

## Currency Substitution

The use of alternative currencies by people is not new, there is an entire body of research devoted understanding this practice, known as *currency substitution*. Currency substitution is defined by @calvo2002 as the highly prevalent use of [foreign]{.underline} fiat currencies to fulfill any of the three functions of money (store of value, means of exchange, unit of account). Note, there is also something known as *official currency substitution*, which is when a government officially adopts a foreign currency as a legal tender in their own country. However, this paper refers to the personal and in unofficial use by individuals.

This research paper draws upon the well-established theories of currency substitution to explore the factors driving cryptocurrency adoption. By examining the similarities between currency substitution and cryptocurrency adoption, this study aims to determine whether the predictors of currency substitution can be built into models of cryptocurrency adoption to improve the explanatory power.

## Literature Gap and Relevance

The research is relevant for both academics and policymakers. The contribution to the academic literature is twofold. Firstly, existing models cannot fully explain the differences in adoption seen across countries. Secondly and more innovatively, the paper provides corresponding analysis for cryptocurrency through the lens of currency substitution, which has not been done before to the best of my knowledge. As will be seen in the section [Dependent Variable: Cryptocurrency Adoption] this research paper makes use of a new and to my knowledge previously not studied panel dataset of Bitcoin adoption that will be able to capture the most recent trends in this area.

In terms of policymakers, it is important for them to be able to understand how changes in underlying economic conditions may influence the use of cryptocurrency as this will have policy implications. It is likely that questions around cryptocurrency will increase in importance in the future due to the increased interest and usage of cryptocurrencies globally: Both from private individuals and governments looking to capitalize on the technology in various ways. Figure \@ref(fig:fig-btc) below shows the trend in cryptocurrency's market capitalization in USD from 2010 - 2025. This increased capitalization is evidence of interest in the technology among market players.

## Research Question

Due to the similarity in potential uses of foreign currency and cryptocurrency, this paper evaluates a model considering not only the predictors of cryptocurrency usage (currency stability, investment, wealth, sins, remittances, capital controls), but also the additional factor coming from the currency substitution literature (sovereign default risk) to see if including this can build an improved model of cryptocurrency adoption. Technology is not explicitly included as a predictor, despite it's discovery in the literature since this is assumed to be fully covered by the introduction of cryptocurrency, itself, a new technology.

The research question can be summarized as: *Do currency stability, investment, wealth, sins, remittances, capital controls and sovereign default risk effect the adoption of cryptocurrency?*

The next section discusses the literature in further detail.

\FloatBarrier

```{r fig-btc, fig.cap="Cryptocurrency Market Capitalization 2010-2025 (Statista, 2025)", echo=FALSE, warning= F, fig.width=6, fig.height=4, fig.pos = 'H'}

# Working Bitcoin Adoption-> 
# Source: https://www.statista.com/statistics/730876/cryptocurrency-maket-value/
path <- "Data/d.Market_Cap_BTC/Statista_Crypto_Market_Cap.xlsx"
d.Market_Cap <- read.xlsx(path,sheet = "Data") 
d.Market_Cap$Date <- as.Date(d.Market_Cap$Date, format="%b %d, %Y")

ggplot(d.Market_Cap, aes(x = Date, y = Capitalization)) +
  geom_line(color = "blue", size = 1) +   # Line plot
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months",
               limits = c(min(d.Market_Cap$Date), max(d.Market_Cap$Date)),
               expand = c(0, 0)) +  # Prevents ggplot from adding extra space
  labs(
       x = "Date",
       y = "Market Cap (in Billion U.S. Dollars)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 6),  # Adjust x-axis text
    panel.grid.major = element_line(size = 0.2, linetype = "solid", color = "gray80"),  # Reduce major gridline thickness
    panel.grid.minor = element_blank()  # Completely remove minor gridlines
  )
```

\FloatBarrier

```{r, echo = F}
# Remove Taiwan (TWN)
d.Adoption <- d.Adoption %>% filter(Country != "TWN")
```

In addition to the increased capitalization, policymakers have begun to carve out a space for blockchain technology in their economies in different ways. Likely the most famous example is El Salvador legislating Bitcoin as a legal tender in their country in 2021 [@bbc2021]. However there are also other examples such as the canton of Zug in Switzerland allowing residents to pay taxes up to CHF 1.5M with certain cryptocurrencies or several large institutional investors and sovereign wealth funds globally purchasing Bitcoin [@chainalysis2024; @kantonzug]. At the time of writing, a recent example is Mubadala Investment's (Abu Dhabi) purchase of USD 436M in Bitcoin [@cryptodaily2025]. All of that is to say, research aiming to understand drivers of cryptocurrency demand is likely to remain relevant or even increase in importance in the foreseeable future.

**Roadmap**

For the definition of the research topic, two questions / fields are of interest: "What drives currency substitution?" and "What drives the adoption of cryptocurrencies?" These topics will now be investigated in separate literature reviews in the next section. From these reviews the research question "[[**RQ HERE**]{.underline}]" seen above was developed. Next, the methods, data and data transformation used to answer the research question will be discussed and visualized. Finally, the statistical results are shown, discussed and placed in the academic context to provide conclusions for researchers and policymakers interested in what factors can drive the adoption of cryptocurrency.

\newpage

# Literature Review

## Currency Substitution

There is a body of academic literature evaluating why individuals use foreign currencies, this section discusses this literature. Those interested in a succinct visual overview should visit Table \@ref(tab:litreviewCS) in the Appendix.

### Currency Stability

The stability of local currencies as a driver for foreign currency adoption is a debated issue in the academic literature, since there are two main ways of looking at currency stability (inflation and exchange rate), these topics are evaluated separate.

#### Inflation

Both perceived and real economic problems are identified in the literature as reasons for people to engaging in currency substitution, the primary economic issue here is inflation. There are quantitative studies, such as those by @vieira2012 and @rennhack2006 finding inflation is a key predictor of currency substitution. Another quantitative paper @honig2009 argues that lack of trust in the stability of the local currency leads people to use foreign currencies. Finally, an implicit argument for the viewpoint that inflation leads to currency substitution is made by @kokenyne2010 who argue countries wishing to stop currency substitution from happening in their domestic economies should focus their efforts on taming excess inflation. This claim is backed up by a practical study of the Turkish economy by @tasseven2015 who argues that foreign currencies were used precisely due to the high and inflation in the 1990s an then began falling out of favor as price stability increased. @levy2021 makes a similar conclusion when he credits inflation first, among a number of factors for the success of many Latin American countries to reduce currency substitution, which is oftentimes seen as negative in the eyes of policymakers. Although a considerable amount of studies credit inflation as positively related to currency substitution, one study focusing on Croatia, Slovenia and Slovakia by @stix2011 did not find inflation to be a contributing factor to currency substitution.

#### Exchange Rate

Another measure of currency stability, the volatility of the exchange rate to major currencies is also oftentimes found to be important in relation to why individuals begin using a foreign currency. Using a threshold ARCH model on 28 countries and an auto regressive distributed lag model of Nigeria, @ju and @ajibola2021 respectively find that there is a correlation between the foreign exchange rate volatility and the use of foreign currencies. Contrary to these findings however, the already mentioned study by @stix2011 also did not find exchange rate volatility to be a contributing factor to currency substitution.

### Risk of Sovereign Default

The risk of a sovereign default also appears in the literature on currency substitution, although less clearly than inflation. @vieira2012 find this to be a stronger predictor of currency substitution than inflation in their quantitative study on 79 economies at different development levels. To the best of my knowledge no other studies have evaluated this convincingly. Although @vieira2012 do quote a number of papers as foundations for evaluating the risk of sovereign in their review, many of these do not clearly claim the logic applied by @vieira2012. Other studies tend in the area focus on official currency substitution and the effect that this choice has on the risk of sovereign default [@berg2000, @sims2001]. Little consideration is given to the case of currency substitution as a choice by individuals and how this choice is influenced by the risk of sovereign default. However, the study be @vieira2012 clearly shows it's importance and therefore this factor should be included in a study on the topic

### Technology

There are also some studies arguing the advancement of technology will aid in facilitating currency substitution. @guidotti1993 argues using a theoretical model, that the reduction on transaction and holding costs to foreign currency, spurred by financial innovation can promote it's use. Such a theory is practically backed up a study of Nigeria by @ujunwa2021 who take an augmented money demand model which includes markers for technology (for example: internet banking transactions) and find that these are strong predictors of foreign currency use. It is therefore not far fetched to argue that a new technology like cryptocurrency could spur the use of currency substitution, even if this implies the use of a new "currency", so long as this currency has the potential to reduce transaction and holding costs, which the fundamentals of certain cryptocurrencies definitely do.

## Adoption of Cryptocurrency

In parallel the the currency substitution literature, there is a wide body of literature studying the usage of cryptocurrency that finds several of the factors connected to currency substitution to be linked to the adoption of cryptocurrency, however most of these studies focus on Bitcoin, rather than other digital assets. Those interested in a succinct visual overview should visit \@ref(tab:litreviewCA) in the Appendix.

### Inflation

There are a number of studies that have evaluated the relationship between cryptocurrency and inflation. @choi2022, @conlon2021 and @gaies2024 study time series data on Bitcoin prices and find that they are correlated positively to inflation or inflation expectations. Mixed evidence is presented by @phochanachan2022 who find the inflation hedge is only present in the short term. @chainalysis2024 evaluated Argentine stablecoin trading data the leading local exchange Bitso and found that with each devaluation against the USD (associated with increase in Argentine inflation), the monthly stablecoin trading volume on Bitso increased, indicating this could be a response to Argentine peso losing it's value. Similar results but for evaluated for Bitcoin instead of stablecoin exist in Venezuela [@chainalysis2024]. Academic case studies of countries using cryptocurrency in response to inflation are limited, only @taskinsoy comes up. He argues that the relative instability of the Turkish Lira is what drives many in the country to use the cryptocurrency Bitcoin instead of the local currency.

Similar studies as mentioned at the beginning of the section study time series data of inflation and the price of Bitcoin, these however find no effect of correlation between Bitcoin [@basher2022; @smales2024]. In studying country economies, both @parino and @ricci2020 find that there is a negative effect of inflation on the price of the cryptocurrency Bitcoin. However it should be noted that the former focused on data from before 2015, which may have been too early to see adoption in developing countries and the latter only evaluated already developed economies which have seen lower levels of inflation compared to developing countries.

### Investment

Investment is found as a key use case for the purchase of cryptocurrencies. @voskobojnikov2020 conduct interviews among North American respondents and find investment is one of the main intended uses of cryptocurrency among non-users. Quantitative studies back this idea up. @glaser2014 find that the pattern of trading on the former Mt. Gox cryptocurrency trading platform implies that users were investing, not using the currency for payments. This is because while the value of currencies on individual accounts did change, the total value on the exchange did not change significantly. To the authors, this suggested that users were shuffling funds between each other, but not using the cryptocurrencies for payments.

### Wealth

Wealth is a well established factor connected to to cryptocurrency in academia, studied through various methods. @lammer2019 studied German bank accounts and found wealthier people were more likely to own Bitcoin. This conclusion is backed up on a national scale by @parino who found GDP per capita to be positively correlated to Bitcoin ownership. Further survey evidence on the average cryptocurrency user is provided by @gemini2021 who find the average cryptocurrency investor has a household income of USD 110K, more than 1.5 times the national average for that year. It should be noted that the survey was not representative and explicit only includes those with a household income above USD 40K, meaning the real average household income of the average investor is likely lower. The results of the studies indicate that volatile assets like cryptocurrency may only be bought only by those who can afford to take temporary losses when the price of the asset decreases.

### Sins

The use of cryptocurrencies in areas that are illegal or considered immoral is also a driver of their usage in many cases. This is a fairly diverse list, so only some illustrative examples will be presented here. It ranges from using Bitcoin to pay for illicit goods and services, such as was done on the now shut down Silk Road dark - web sites [@saurabh2017]. Research has found that countries with larger shadow economies, the Bitcoin trading volume is much more strongly responsive to shocks to the shadow marked (raids, seizures), indicating the cryptocurrency Bitcoin is used for illicit transactions [@marmora2021].

Cryptocurrencies may also be used extensively by sanctioned countries to settle international debts. Iranian academics have evaluated the possibility of using cryptocurrencies to settle debts for their resource exports, which is difficult to do in the current international settlement architecture given the punitive international sanctions placed on the country [@sarvi2020]. Venezuela developed the *Petro*, a coin who's value was tied to the country's oil reserves and was meant to serve as a means of settling international payments for the sanctioned country, the project was discontinued in 2024 [@rfi2024].

Sanctions on individuals also seem to b a driver of adoption of cryptocurrencies. According to an analysis by @chainalysis2020 75% of all cryptocurrency transaction on a randomly selected Venezuelan exchange were over USD 1K, given the relatively low wages in the country, it is likely that this represents sanctioned individuals attempting to move funds out of the country. Similarly, Russian language cryptocurrency exchanges without Know Your Customer regulations have tripled since the start of the war in Ukraine and the following sanctions on both individuals connected to the Kremlin and ordinary Russians. The idea that this is connected to the sanctions is supported by the abscence of growth in Russian speaking exchanges that do conduct Know Your Customer checks and presumably comply with international sanctions [@chainalysis2024]. Further evidence is provided by @alnasaa2022 who see higher adoption of Bitcoin in corrupter countries, indicating the possibility that corrupt officials are using the cryptocurrency Bitcoin to move proceeds from corruption.

### Remittances

Another potential reason for the adoption of cryptocurrency is for remittance payment. This has not been studied extensively academically but the economic fundamentals and some practical examples show the potential. Fees for remittance payments can be very expensive, between 6.9-20% according to @ruehmann2020. Simultaneously blockchain technology can have incredibly low fees, typically between 0-1% according to @dyhrberg2018. This low cost has led some academics like @folkinshteyn2015 to argue cryptocurrencies like Bitcoin could form an important aspect of lowering remittance costs. This cost advantage, was the official reason behind El Salvador making Bitcoin legal tender in 2021 [@bbc2021]. Data from other areas and tokens supports the cost idea as well, @chainalysis2024's data suggests 60% lower costs for a USD 200 payment in sub - Saharan Africa when using a transfer system based on stablecoins compared to a traditional fiat supported system. In addition, there was at one point a concerted effort by the Libra Association (Facebook / Meta) to release a stablecoin that was to be integrated with existing and widely used Meta communications platforms such as Whatsapp. Through this integration, it had the potential to reach the 1.1B people globally who have a mobile phone, but no bank account [@worldbank2018]. While challenges like internet access and identity verification for users would have remained, the potential of this stablecoin integrated in communication service for remittances hard to deny [@ruchti]. Ultimately, the project ended due to regulatory opposition from the United States [@mcnickel2024]. All of that is to say that the potential for cryptocurrencies or the blockchain architecture to increase financial inclusion is there and has been explored.

### Capital Controls

There exists research that claims capital controls to be relevant to the adoption of cryptocurrency. @carlson2016 conducts expert interviews on the Argentine example and finds that capital controls can and are being evaded using Bitcoin. @hu2021 study Chinese Bitcoin transaction and conclude that 25% of the transaction volume represents capital flight out of the country. @viglione2015 find a similar result in a quantitative analysis of multiple economies. They see a "premium" being paid for Bitcoin in these countries, which they interpret as an extra demand for Bitcoin relative to other countries, which they interpret as "extra demand". Additional evidence for the importance of capital controls is provided by @alnasaa2022 who run a cross country analysis including capital controls as a predictor and find the capital controls to be a statistically significant predictor of cryptocurrency usage. The study of capital controls as a predictor in any field is limited by the diversity of potential measures to restrict capital flow and the lack of a standardized metric.[^2]

[^2]: Note: An index such as the one produced in this paper from regularly published IMF Data could form the basis for consistent and replicable study of capital controls. See Section \@ref(predictors-independent-and-control-variables).

\newpage

# Methodology

This section discusses the quantitative methods to answer the research question. It is organized as follows. Firstly, common terms among the models are introduced, then the formula defining fitted input - output relationship between the independent and dependent variables are described, with additional terms being defined where necessary. Next, the hypothesis and significance level are discussed. Finally, benefits and drawbacks are discussed in relation to how well the methodology can answer the research question.

## Common Terms

The models have a number of shared terms. These are listed and explained below.

$\beta_{1,2...7}$: Coefficients for each independent variable. The exact interpretation of these depends on the input-output relationship specified in the models, but they all describe a certain change in the dependent variable connected to a change in the input variable.

$i$: Denotes the cross-sectional unit (country) in the panel data.

$t$: Denotes the time unit (year)

$\varepsilon$: The error term, capturing unobserved factors

## Model 1: Linear Regression - No Transformation

The first model is a linear regression with the temporal aspect modelled as a categorical. A statistically significant term for the year dummy will have no effect on answering the research question. The formula can be seen below. The results for this model can be seen in the section \@ref(model-1-linear-regression-no-transformation).

```{=tex}
\begin{align*} 
\text{Cryptocurrency Adoption}_{i,t} &= \beta_0 + \sum_{t=2020}^{2023} \beta_t \cdot D_{t,i}\\ &+ (\beta_1 \cdot \text{Currency Stability}_{i,t}) \\ 
&+ (\beta_2 \cdot \text{Investment}_{i,t}) \\ 
&+ (\beta_3 \cdot \text{Wealth}_{i,t}) \\
&+ (\beta_4 \cdot \text{Sins}_{i,t}) \\
&+ (\beta_5 \cdot \text{Remittances}_{i,t}) \\
&+ (\beta_6 \cdot \text{Capital Controls}_{i,t}) \\
&+ (\beta_7 \cdot \text{Sovereign Default Risk}_{i,t}) + \varepsilon_{i,t} 
\end{align*}
```
The group of terms $\sum_{t=2020}^{2023} \beta_t \cdot D_{t,i}$ represent the categorical variables of each year with the first year as the baseline. The dummy $D_t,i$ is 1 when the corresponding year is "present" and 0 otherwise. It ensures that for each year only the relevant categorical shift in year is performed as only the one relevant year's term will be counted towards the summation. The summation starts at the year 2020, instead of the first year 2019 in the dataset, because 2019 is the baseline.

$\beta_0$: The intercept, representing the baseline level of cryptocurrency adoption when all continuous predictors are equal to zero and the year is at the 2019 baseline.

The terms $\beta_{1,2...7}$ represent the response in dependent variable as a result of increasing the corresponding variable by 1 while holding all of the other variables constant.

## Model 2: Linear Regression - Transformations

The second model extends the previous model by explicitly transforming certain variables. This is done to better meet the assumptions of a linear regression, which requires normally distributed variables. As will be seen in the section \@ref(exploratory-data-analysis), not all of the variables clearly meet this assumption. It is therefore worth to run a model with transformations applied. It should be noted that the interpretation of the coefficients will become altered as a result of this.

The transformation used will be the **Yeo - Johnson** transformation, which is an extension of the Box - Cox transformation allowing both negative and 0 numbers, which there are in the data studied. Such transformations can be beneficial in cases where the distribution assumptions of the input data is not met.

The Yeo - Johnson transformation makes the following transformation by estimating a $\lambda$ value from the data. The exact values for $\lambda$ used for the variables can be seen in Table \@ref(tab:YJ-lambda-tbl) in section \@ref(data-preparation).

```{=tex}
\begin{align*}   
\psi(X, \lambda) =
\begin{cases} 
\frac{(X + 1)^\lambda - 1}{\lambda} & \text{if } \lambda \neq 0, X \geq 0 \\[10pt]
\log(X + 1) & \text{if } \lambda = 0, X \geq 0 \\[10pt]
\frac{- \left[(-X + 1)^{2 - \lambda} - 1\right]}{2 - \lambda} & \text{if } \lambda \neq 2, X < 0 \\[10pt]
-\log(-X + 1) & \text{if } \lambda = 2, X < 0
\end{cases}
\end{align*}
```
This means that the formula for the transformed linear regression becomes the following:

```{=tex}
\begin{align*} 
\psi (\text{Cryptocurrency Adoption}_{i,t}, \lambda) &= \beta_0 + \sum_{t=2020}^{2023} \beta_t \cdot D_{t,i} \\
&\quad + \beta_1 \cdot \psi (\text{Currency Stability}_{i,t}, \lambda) \\ 
&\quad + \beta_2 \cdot \psi (\text{Investment}_{i,t}, \lambda) \\ 
&\quad + \beta_3 \cdot \psi (\text{Wealth}_{i,t}, \lambda) \\
&\quad + \beta_4 \cdot \psi (\text{Sins}_{i,t}, \lambda) \\
&\quad + \beta_5 \cdot \psi (\text{Remittances}_{i,t}, \lambda) \\
&\quad + \beta_6 \cdot \psi (\text{Capital Controls}_{i,t}, \lambda) \\
&\quad + \beta_7 \cdot \psi (\text{Sovereign Default Risk}_{i,t}, \lambda) + \varepsilon_{i,t}
\end{align*}
```
The term $B_t$ represents **[[add here]{.underline}]**.

The terms $B_{1,2...7}$ represent **[[add here]{.underline}]**.

## Model 3: Fixed Effects

Panel data gives the choice between fixed or random effects to account for omitted variable bias [@dougherty2011]. In practical applications the choice of fixed or random effect is decided using the **Hausman Test**. The results of this test can be seen in Table \@ref(tab:ht-tbl) in section \@ref(hausman-test). The Hausman Test indicates that a Fixed Effects Model should be used. The formula can be seen below.

```{=tex}
\begin{align*}  
\text{Here Specification for Model 3}
\end{align*}
```
[[**Here describe any neccesary model specifications**]{.underline}]

## Hypothesis

Formally, the null and alternative hypothesis, in words and mathematically are:

**Null Hypothesis**

$H_0$: Inflation, investment, wealth, sins, remittances, capital controls and\newline sovereign default risk have no statistically significant effect on cryptocurrency adoption.\newline \newline$H_0:\beta_1=\beta_2=\beta_3=\beta_4=\beta_5=\beta_6=\beta_7=0$

**Alternative Hypothesis**

$H_1$: At least one independent variable has a statistically significant\newline effect on cryptocurrency adoption.\newline\newline $H_1: \exists \beta_j \neq 0, \quad \text{for at least one } j \in \{1,2,3,4,5,6,7\}$

**Direction of Effect**

The alternative hypothesis does not specify a direction of effect due to the limited and conflicting research on both cryptocurrency adoption and currency substitution.

**Significance Level**

Due to the limited data size and therefore statistical power of this paper (see [Underlying Data]) a significance level at the upper end of the normal range ($\alpha=0.1$) will be used.

## Internal Validity

-   Can we draw cause and effects?

-   Strength of Proxies?

## External Validity

Due to the broad range of countries included in this study, the generalizability of the results should be broad. Figure \@ref(fig:fig-worldmap-selected) shows the countries available in the @statista_adoption dataset and being studied explicitly in this paper. It is clear that a diversity of countries are represented by the data, in different socioeconomic aspects such as those found as relevant in the literature review for both currency substitution and cryptocurrency adoption. The main criticism in terms of generalizability would likely be the under inclusion of African countries, with just 4 out of the 54 countries on the continent represented. Nevertheless, the the argument can be made that this research will be applicable to most countries that who's economic data falls within the range of the independent variables. Extreme political and economic outliers such as North Korea will not fall within this scope, but that is typical of almost any national level panel data analysis.

\FloatBarrier

```{r fig-worldmap-selected, echo = F, fig.width=6, fig.height=4, fig.pos = 'H', fig.cap= "Map Showing Countries With Available Cryptocurrency Adoption Data"}
world <- ne_countries(scale = "medium", returnclass = "sf")
display_country_codes <- d.Adoption$Country
world <- world %>%
  mutate(highlight = ifelse(iso_a3 %in% display_country_codes, "Yes", "No"))
ggplot(data = world) +
  geom_sf(aes(fill = highlight), color = "black", size = 0.01) +
  scale_fill_manual(values = c("Yes" = "blue", "No" = "lightgray")) +
  theme_minimal() +
  labs(fill = "Data Availability?")

```

\FloatBarrier

\newpage

# Underlying Data

This section discusses the underlying data used for this research paper. It should be noted that for some of the data presented here, the country variable is already altered using the library {countrycode}, which can take different versions of country names and obtain the world bank code (ISO 3), this helps with correctly and repeatbly joining the data later on.

## Dependent Variable: Cryptocurrency Adoption

Cryptocurrency adoption is measured using a 2024 dataset by @statista_adoption. The dataset is the results of a survey where respondents were asked if they had used cryptocurrency in the given year. The data's first few rows can be seen in Table \@ref(tab:Adoption) below. Due to the unavailability of data for the other variables for the year 2024, only data up to 2023 will be used from this set. The data is available for 56 countries of different levels of economic development.

The data quality of the survey leaves little room practical room for improvement. While survey was not representative and voluntary, opening the data quality up to self - selection issues, the only real improvement could have been a mandatory national census. The authors describe their approach to data collection and it is clear and professional: sample sizes of at least 2k people per country, checks for bots and speed racers through in the survey and the survey was performed in the official languages of the country [@statista2020]. That is to say, the quality of this variable is sufficient.

[[**Does "Cryptocurrency" as asked in the survey actually reflect the same kind of cryptocurrency I am looking at here**]{.underline}**?**]

```{r Adoption, echo = F }
# Display table
d.Adoption %>%
  head() %>%
  kable(col.names = c("Country", "2019", "2020", "2021", "2022", "2023", "2024"), 
        caption = "Statista (2024a) Data: \\ Share (Percentage) of Respondents Who Reported Using Cryptocurrency in Select Years") %>%
  kable_styling(full_width = FALSE, latex_options = "hold_position")
```

## Predictors: Independent and Control Variables

Table \@ref(tab:datatable) shows an overview of the indicators used as Proxies and their sources. The indicators for currency stability, wealth, remittances and risk of sovereign default are self-explanatory and should be of sufficient quality as they are sourced primarily from the World Bank, which is a state of the art data source for cross country economic data analysis.

```{r, datatbl, echo = F, results = "asis"}
# Create the data frame with citations and special characters handled
lit_data_sources <- data.frame(
  Indicator = c( "Inflation, consumer prices (annual %)",
                 "Gross domestic savings (% of GDP)", 
                 "GDP per capita (current US$)",
                 "Personal remittances, received (% of GDP)",
                 "External Debt (% of GDP)",
                 "Political Corruption Index (see below)",
                 "Bespoke Capital Controls Index (see below)"
                 #"Mobile Access (%)",
                 ),
  Proxy_for = c("Currency Stability", 
                "Investment", 
                "Wealth",
                "Remittances",
                "Risk of Sovereign Default",
                "Sins",
                "Capital Controls"
                #"Technology",
                ),
  Source = c("World Bank (2024c)",
             "World Bank (2024b)",
             "World Bank (2024a)",
             "World Bank (2024d)",
             "Focus Economics (2024)",
             "V-Dem (2024)",
             "IMF (2024)"
             #"World Bank",
             
  )
)

# Generate the table using kable
lit_data_sources %>%
  kable(col.names = c("Indicator", "Proxy for", "(Primary) Source"), caption = "Overview of Data Sources for Independent Variables (\\#tab:datatable)") %>%
  kable_styling(full_width = F, latex_options = "hold_position")


```

The following proxies must be discussed in further detail: Investment, sins, capital controls.

**Investment**

Investment is related to Savings in National Accounting, more closely in some, than in other models. In the classical view of a closed economy without government spending, savings are equal to planned investment [@mitchell2019]. This make national savings as a percentage of GDP a good proxy for available investment funds.

**Sins**

A single indicator is used to encompass all of the "sin" uses of Bitcoin. The two primary sinful uses are criminality and the evasion of sanctions. Since Western countries routinely sanction individuals and not the countries themselves based on corruption, human rights abuses and other serious accusations, it makes sense to use corruption as a proxy for individual sanctions that people may attempt to circumvent using Bitcoin [@u.s.departmentofthetreasury2022]. Using corruption as a proxy for crime is also a possible approach as the link between corruption and (in particular organized) crime as been shown several regions and studies [@buscaglia2003; @centerforthestudyofdemocracy2010; @mazzitelli2007]

Therefore, the political corruption index, published by @politica2024 is used in an attempt to cover both crime and the likelihood that individuals attempt to move illegally obtained money abroad. The index is made up of several subsets of corruption. In a study with more data quantity available it would make sense to use specific types of corruption from this index combined with other indicators for criminality. However, due to the already small data size, the trade off of including several variables for the sinning attribute identified in the literature would be too adverse on the statistical power. Therefore, the aggregated corruption index is used, rather than one of the specific corruption indicators, like for example executive branch corruption [@olin]. The @politica2024 data can be downloaded directly via library {vdemdata} in R-Studio, after connecting to GitHub. The data is available only up to and including 2023.

```{r, warning = F, warning = F, message = F, echo = F }
# install.packages("devtools")
# uncomment above lines when running for the first time
devtools::install_github("vdeminstitute/vdemdata")
library(vdemdata)
d.Corruption <- vdem[, c("country_name", "year", "v2x_corr")] # relevant variables
d.Corruption <- d.Corruption %>%
  filter(year %in% c(2019: 2024)) # relevant years
d.Corruption$Country <- countrycode(d.Corruption$country_name, "country.name", "iso3c")
# above: world bank / ISO 3 code added
```

```{r, echo = F}
# Renaming, dropping and re-arranging column
d.Corruption <- d.Corruption %>%
  rename(Corruption = last_col(offset = 1)) %>%
  select(Country, everything(), -country_name) 

# Dropping Gaza / Westbank / NA (represents Kosovo, Somaliland and Zanzibar: These are causing aggregation issues and are not in the depdendent variable so can be safely removed
d.Corruption <- d.Corruption %>%
  filter(!Country %in% c("PSE", NA))


# Converting to Wide Format Panel Data
d.Corruption <- d.Corruption %>%
 pivot_wider(names_from = year, values_from = Corruption)  #%>% 
 # mutate(`2024` = NA)
```

**Capital Controls**

Since capital controls have been identified as important in section [Capital Controls], they must be accounted for an a model attempting to explain cryptocurrency adoption. Unfortunately there is a lack of structured and recent data around this topic. The most recent dataset was created by @fernandez2016 and was updated with data up to 2017, who produced an index to measure the severity of capital controls. The source used here will be from the online query tool of @imf2024 which allows the recovering of information contained in the annually published Report on Exchange Arrangements and Exchange Restrictions, specifically the 5 indicators: "Controls on Personal Payments", "Prior Approval", "Quantitative Limits", "Indicative Limits / Bona Fide Test" and "Controls on Personal Capital Transactions". The key limitation of this dataset is that the data is only available up to and including 2022. Techniques to deal with missing data will be used to make this dataset usable, since to my knowledge it is the best available way to account for capital controls in a consistent and repeatable way. The first rows of the raw data can be seen in Table \@ref(tab:d).

```{r, d, echo = F}
path <- "Data/d.CC/IMF (2024).xlsx"
d.CC <- read.xlsx(path)
d.CC %>%
  head() %>%
  kable(
    col.names = c("Year", "IFS Code", "Country", 
                  "Controls Personal \\ Payments",  
                  "Prior \\ Approval",
                  "Quantitative \\ Limits",
                  "Indicative Limits / \\ Bona Fide Test",
                  "Controls on Personal \\ Capital Transactions"),
    escape = FALSE,  # Allows LaTeX commands like \\ to be used
    caption = "IMF (2024) Capital Controls Dummy Data",
    format = "latex"
  ) %>%
  kable_styling(full_width = F, latex_options = c("scale_down","H")) %>%
  column_spec(1, width = "0.7cm") %>%  # Custom width for Year
  column_spec(2, width = "1.5cm") %>%    # Custom width for IFS Code
  column_spec(3, width = "1.7cm") %>%    # Custom width for Country
  column_spec(4, width = "2.2cm") %>%    # Custom width for Controls Personal Payments
  column_spec(5, width = "2.0cm") %>%  # Custom width for Prior Approval
  column_spec(6, width = "2.3cm") %>%  # Custom width for Quantitative Limits
  column_spec(7, width = "3.3cm") %>%  # Custom width for Indicative Limits / Bona Fide Test
  column_spec(8, width = "3.5cm")      # Custom width for Controls on Personal Capital Transactions
```

In order to turn this into a quantitative variable, representing the strength of capital controls, these variables will be turned into an an index from 0 - 1 by assigning a value of 1 for each "yes" and 0 for each "no" and then dividing the result by the number of available data points for that country in that year, conceptually, this means that an index calculated with just one available data point look the same as one with all data points available. Additionally, it should be noted, by using an equally weighted index generating method, the assumption is made that each of these types of restrictions is equally important in the types of capital controls that influence the adoption of cryptocurrency. In the case that a country and year has no data point (only 1 example in the data), a NA is assigned to this value. Table \@ref(tab:tbl-ind-head-CC) shows the top rows of the resulting dataframe containing the index.

```{r d.CC-index-generation, echo = F,warning = F}
# Rename the last 5 columns to A, B, C, D, E
colnames(d.CC)[(ncol(d.CC)-4):ncol(d.CC)] <- c("A", "B", "C", "D", "E")

# Compute Index dynamically based on available (non-NA) values
d.CC$IndexCC <- rowSums(d.CC[, c("A", "B", "C", "D", "E")] == d.CC[11,4], na.rm = TRUE) / 
                 rowSums(!is.na(d.CC[, c("A", "B", "C", "D", "E")]), na.rm = TRUE)

d.CC$IndexCC[is.nan(d.CC$IndexCC)] <- NA # #converting 0 data point to NA

d.CC$Country <- countrycode(d.CC$Country, "country.name", "iso3c")

d.CC <- na.omit(d.CC, cols = "Country")
# Above Warnings Countries are not found in the Adoption set, can be safely removed

d.CC <- d.CC %>% select(Country, Year, IndexCC) # rearrannging and removing rows


d.CC <- d.CC %>%
  pivot_wider(names_from = Year, values_from = IndexCC) %>%  # make wide format
  mutate(`2023` = NA)#, `2024` = NA)
```

```{r tbl-ind-head-CC, echo = F}
kable(head(d.CC), caption = "Head of Table with Capital Controls Index",
      row.names = F)
```

## Data Preparation

This section discusses how the data is treated to prepare it for analysis.

### Removing Countries not in *d.Adoption*

Countries without a dependent variable are removed in this step, since without a dependent variable nothing of insight can be gained, even if all the dependent variables are there. Due to lacking independent variable data, the value for Taiwan is also removed.

```{r removing-irrelevant-countries, echo = F, message = F, warning = F, include = F }
# Extract unique country names from d.Adoption
relevant_countries <- unique(d.Adoption$Country)

# List of original dataframes
IV_data <- list(
  d.CC = d.CC, 
  d.Corruption = d.Corruption, 
  d.ED = d.ED, 
  d.GDP = d.GDP, 
  d.GDS = d.GDS, 
  d.Inflation = d.Inflation, 
  d.RR = d.RR
)

# Find missing countries for each dataset
IV_countries <- lapply(IV_data, function(df) unique(df$Country))
missing_countries <- lapply(IV_countries, function(countries) setdiff(relevant_countries, countries))

# Print missing countries for each dataset
for (name in names(missing_countries)) {
  cat("\nMissing countries in", name, ":\n")
  print(missing_countries[[name]])
}

# Find countries that are missing in **any** dataset
all_missing_countries <- unique(unlist(missing_countries))

cat("\nCountries in d.Adoption but missing in at least one dataset:\n")
print(all_missing_countries)

# Apply filtering and overwrite original dataset names
filtered_datasets <- lapply(IV_data, function(df) {
  df[df$Country %in% relevant_countries, , drop = FALSE]
})

# Assign filtered datasets back to their original names in the environment
list2env(filtered_datasets, envir = .GlobalEnv)
```

### Removing Year 2024 from *d.Adoption*

Since there is no data for any of the other indicators for the year 2024 in a structured and accessible format, the year 2024 is removed from *d.Adoption*.

```{r removing 2024, echo = F}
d.Adoption <- d.Adoption %>% select(-'2024')
```

### Manually Adding Missing Data

Due to the already small data size, where possible, missing independent variables are filled in manually. The following information was added in this manner. While this approach is not consistent throughout and therefore limits practical replicability, it offers the benefit of retaining slightly increased data quantity.

-   Argentina Inflation Rate 2018-2022 [@statistaArgentinaInflation]

-   Nigeria Gross Domestic Savings (% of GDP) 2018-2021 [@nigeria]

-   Belgium, Canada, France, Ireland, Spain External Debt (% of GDP) 2018-2023 [@ceicdata2025]

```{r manual-impute-argentina, echo = F}
NGA_GDS <- c("2019" = 19.83, "2020" = 27.38, "2021" = 32.73, "2022" = NA, "2023" = NA)#, "2024" = NA)

NGA_row <- which(d.GDS$Country == "NGA")
d.GDS[NGA_row, 2:ncol(d.GDS)] <- NGA_GDS  # Assuming the first column is "Country"

Arg_Inf <- c("2019" = 53.55, "2020" = 42.02, "2021" = 48.41, 
             "2022" = 72.43, "2023" = 133.49)#, "2024" = NA)

arg_row <- which(d.Inflation$Country == "ARG")
d.Inflation[arg_row, 2:ncol(d.Inflation)] <- Arg_Inf  # Assuming the first column is "Country"

new_countries_External_Debt <- tibble(
  Country = c("BEL", "CAN", "FRA", "IRL", "ESP"),
  `2019` = c(257.579, 134.234, 235.061, 742.138, 169.725),
  `2020` = c(268.416, 149.077, 265.191, 702.699, 200.181),
  `2021` = c(260.193, 143.214, 258.592, 677.751, 193.562),
  `2022` = c(237.91, 134.904, 244.258, 577.158, 172.805),
  `2023` = c(237.68, 143.661, 245.047, 563.897, 165.481)#,
  #`2024` = c(NA, NA, NA, NA, NA)
)

# Add only missing countries
existing_countries <- d.ED$Country
missing_countries <- new_countries_External_Debt %>% filter(!Country %in% existing_countries)

# Append missing countries to d.GDS
d.ED <- bind_rows(d.ED, missing_countries)
```

### Missing Data Structure

Since not all data can be added manually due to lacking reliable and consistent data sources, alternative techniques must be employed. The type of way that missing values are dealt with depends on the quantity and structure of missing data. A standard approach is to say that any dataset having less than 5% missing can be treated with univariate imputation, such as adding the mean of a row or column for a missing piece of data. However, the structure of the missing data plays a role too. For this reason a Missing Completely at Random (MCAR) is performed [@schwarz2024]. Table \@ref(tab:tblMCAR) below shows the result of this test. The results indicate that univariate imputation can be used for all variables since the hypothesis, that the data is not MCAR is either rejected at the 5% level or the number of missing values is less than 5%. The percentage missing applies only to the numeric columns, so the inclusion of country names in the table do not downward skew the proportion of missing numbers.

```{r fig-MCAR-fun, echo = F, message = F, warning=F}

# Function to run MCAR tests and format results with improved headers
run_mcar_tests <- function(...) {
  dfs <- list(...)
  df_names <- sapply(substitute(list(...))[-1], deparse)  # Get dataframe names
  
  results <- data.frame(
    `Dataset` = character(),
    `Test Statistic` = character(),  # More descriptive header
    `Degrees of Freedom` = character(),
    `P-Value` = character(),  # Improved readability
    `Missing Patterns` = character(),
    `Missing Percent` = character(),  # More readable title
    `Significance Level` = character(),  # Full description
    stringsAsFactors = FALSE
  )

  for (i in seq_along(dfs)) {
    df <- dfs[[i]]
    df_name <- df_names[i]

    # Calculate missing percentage only for numeric fields
    numeric_cols <- df[, sapply(df, is.numeric), drop = FALSE]  # Select numeric columns
    total_numeric_values <- length(numeric_cols) * nrow(df)  # Total numeric data points
    total_missing_values <- sum(is.na(numeric_cols))  # Count missing values in numeric columns
    missing_percent <- round((total_missing_values / total_numeric_values) * 100, 2)

    # Check for missing values
    if (sum(is.na(df)) == 0) {
      results <- rbind(results, data.frame(
        `Dataset` = df_name,
        `Test Statistic` = "-",
        `Degrees of Freedom` = "-\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0No",
        `P-Value` = "Missing",
        `Missing Patterns` = "Values",
        `Missing Percent` = "0.00",  # No missing values
        `Significance Level` = "-"
      ))
    } else {
      test_result <- naniar::mcar_test(df)
      test_result <- as.data.frame(test_result)
      p_value <- test_result$p.value[1]

      # Determine significance level
      significance <- case_when(
        p_value < 0.01  ~ "Highly Significant",
        p_value < 0.05  ~ "Significant",
        TRUE            ~ "Not significant"
      )
      
      results <- rbind(results, data.frame(
        `Dataset` = df_name,
        `Test Statistic` = round(test_result$statistic[1], 2),
        `Degrees of Freedom` = test_result$df[1],
        `P-Value` = round(p_value, 4),
        `Missing Patterns` = test_result$missing.patterns[1],
        `Missing Percent` = paste0(missing_percent),
        `Significance Level` = significance
      ))
    }
  }
  
  # Print results in a clean table using kable
  colnames(results) <- c("Dataset", "Test Statistic", "Degrees of Freedom", 
                       "P-Value", "Missing Patterns", "Proportion Missing (%)", 
                       "Significance Level")
  
  kable(results, caption = "MCAR Test Results", digits = 4, format = "markdown")
}

library(dplyr)
library(naniar)
library(knitr)
library(kableExtra)

run_mcar_tests1 <- function(...) {
  dfs <- list(...)
  df_names <- sapply(substitute(list(...))[-1], deparse)  # Get dataframe names
  
  results <- data.frame(
    `Dataset` = character(),
    `Test Statistic` = character(),  # More descriptive header
    `Degrees of Freedom` = character(),
    `P-Value` = character(),  # Improved readability
    `Missing Patterns` = character(),
    `Missing Percent` = character(),  # More readable title
    `Significance Level` = character(),  # Full description
    stringsAsFactors = FALSE
  )

  for (i in seq_along(dfs)) {
    df <- dfs[[i]]
    df_name <- df_names[i]

    # Calculate missing percentage only for numeric fields
    numeric_cols <- df[, sapply(df, is.numeric), drop = FALSE]  # Select numeric columns
    total_numeric_values <- length(numeric_cols) * nrow(df)  # Total numeric data points
    total_missing_values <- sum(is.na(numeric_cols))  # Count missing values in numeric columns
    missing_percent <- round((total_missing_values / total_numeric_values) * 100, 2)

    # Check for missing values
    if (sum(is.na(df)) == 0) {
      results <- rbind(results, data.frame(
        `Dataset` = df_name,
        `Test Statistic` = "-",
        `Degrees of Freedom` = "-\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0\u00A0No",
        `P-Value` = "Missing",
        `Missing Patterns` = "Values",
        `Missing Percent` = "0.00",  # No missing values
        `Significance Level` = "-"
      ))
    } else {
      test_result <- naniar::mcar_test(df)
      test_result <- as.data.frame(test_result)
      p_value <- test_result$p.value[1]

      # Determine significance level
      significance <- case_when(
        p_value < 0.01  ~ "Highly Significant",
        p_value < 0.05  ~ "Significant",
        TRUE            ~ "Not significant"
      )
      
      results <- rbind(results, data.frame(
        `Dataset` = df_name,
        `Test Statistic` = round(test_result$statistic[1], 2),
        `Degrees of Freedom` = test_result$df[1],
        `P-Value` = round(p_value, 4),
        `Missing Patterns` = test_result$missing.patterns[1],
        `Missing Percent` = paste0(missing_percent),
        `Significance Level` = significance
      ))
    }
  }
  
  # Rename columns for better readability
  colnames(results) <- c("Dataset", "Test Statistic", "Degrees of Freedom", 
                         "P-Value", "Missing Patterns", "Proportion Missing (%)", 
                         "Significance Level")

  # Print results using kable with column width adjustments
  kable(results, caption = "MCAR Test Results", digits = 4, format = "latex", booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, latex_options = c("hold_position", "scale_down")) %>%
  column_spec(6, width = "4cm") %>%  # Slightly reduce "Proportion Missing (%)"
  column_spec(7, width = "5cm")  # Keep "Significance Level" but not too large
}

```

```{r tblMCAR, echo = F, fig.cap= "MCAR Test on Datasets (excluding d.CC)", warning=F}
run_mcar_tests1(d.Adoption, d.Inflation, d.GDS, d.GDP, d.RR, d.ED)
```

The MCAR test could not be performed for the capital controls data (*d.CC*) as there is an entire column missing, as can be seen in Figure \@ref(fig:md-dCC), which is incompatible with the algorithm of the MCAR test. For this reason, in the interest of maintaining the highest reasonable data quantity, the 2022 (most recent) value of the *d.CC* will be imputed in for the 2023 value of *d.CC*. For a detailed guide on the interpretation of the MD Pattern figure, please see \@ref(interpretation-of-missing-data-pattern-figure). The logic behind this imputation is twofold. Firstly, it is unlikely for capital controls to significantly change in a single year, as it would be require a stark policy shift. Secondly, this type of imputation should be applied here since (mostly) complete data is available for the other indicators for 2023.

```{r md-dCC, echo = F, fig.cap = "Missing Pattern of d.CC", fig.width=5, fig.height=4, message= F,  results="hide"}
md.pattern(d.CC[,-1])
```

```{r imputing dcc, echo = F, message= F, fig.width=5, fig.height=4,  results="hide"}
d.CC[[ncol(d.CC)]] <- d.CC[[ncol(d.CC) - 1]]
remove(filtered_datasets,IV_countries,IV_data,lit_data_sources,world,new_countries_External_Debt)
```

### Imputing Missing Data

Since the missing data is of a structure which can use univariate imputation as shown by the MCAR test, this is done. The assumption behind the way that the imputations are done is that the country in which an observation happens is more important than the year in which it happens. This it is preferred to impute using the country's available data over the year's available data. Due to this dataset taking place during the Covid-19 pandemic, it would not be suitable to just take a mean of all available years for a country. Instead, the average of the year before and after the missing data is used. If a datapoint is missing in the last year, only the previous [available]{.underline} year's data from that country is used. If a data point is missing in the first year, only the first [available]{.underline} year's data point of that country is used to impute. As can be seen in Figures \@ref(fig:md-Adoption) through \@ref(fig:md-ED-GDP) in section \@ref(appendix-2-missing-data-patterns), there are many cases where the missing pattern is not in between available data points and will rely on a single number for imputation. For the sake of brevity, this process is referred to as Nearest Average Imputation.

To be exact, the following imputations imputations are performed:

-   *d.Adoption* completed using Nearest Average Imputation.

-   *d.Inflation* completed using Nearest Average Imputation.

-   *d.GDS* completed using Nearest Average Imputation.

-   No imputations required for *d.GDP* (complete data).

-   No imputations required for *d.Corruption* (complete data).

-   Average for all observations, separated by year, imputed for 1 country value in *d.RR*, as there were no values for any year for that country (United Arab Emirates). No other imputations had to be applied.

-   No imputations required for *d.CC.*

-   No imputations required for *d.ED.*

With those changes, the data is ready for analysis. Please note, practically all datasets are ran through the algorithm to perform the Nearest Average Imputation, this is done to increase replicability of the underlying script.

```{r imputation-fun, echo = F}
fill_na_with_nearest_avg <- function(df) {
  # Compute column means (excluding 'Country' column), ignoring NA values
  col_means <- colMeans(df[, -1], na.rm = TRUE)

  for (i in 1:nrow(df)) {
    row_values <- df[i, 2:ncol(df)]  # Extract numeric part of the row (excluding country column)

    if (all(is.na(row_values))) {
      # If all values are NA, replace with column means
      df[i, 2:ncol(df)] <- col_means
    } else {
      for (j in 2:ncol(df)) { # Start from 2nd column to avoid modifying 'Country'
        if (is.na(df[i, j])) {
          prev_value <- NA
          next_value <- NA
          
          # Find previous non-NA value safely
          if (j > 2) {
            prev_values <- na.omit(as.numeric(df[i, 2:(j-1)]))  # Convert to numeric vector
            if (length(prev_values) > 0) {
              prev_value <- prev_values[length(prev_values)]  # Last previous value
            }
          }
          
          # Find next non-NA value safely
          if (j < ncol(df)) {
            next_values <- na.omit(as.numeric(df[i, (j+1):ncol(df)]))  # Convert to numeric vector
            if (length(next_values) > 0) {
              next_value <- next_values[1]  # First next value
            }
          }

          # Ensure `prev_value` and `next_value` are single numeric values
          prev_exists <- !is.na(prev_value) && is.numeric(prev_value)
          next_exists <- !is.na(next_value) && is.numeric(next_value)

          # Fill NA using the nearest available values
          if (prev_exists && next_exists) {
            df[i, j] <- (prev_value + next_value) / 2  # Average of both nearest values
          } else if (prev_exists) {
            df[i, j] <- prev_value  # Use only the previous value
          } else if (next_exists) {
            df[i, j] <- next_value  # Use only the next value
          }
        }
      }
    }
  }
  return(df)
}

# Example usage:
# df <- fill_na_with_nearest_avg(df)

```

```{r performing impuations, echo = F}
# For repeatability, all datasets are still run through the algorithm
d.Adoption_imp <- fill_na_with_nearest_avg(d.Adoption)
d.Inflation_imp <- fill_na_with_nearest_avg(d.Inflation)
d.GDS_imp <- fill_na_with_nearest_avg(d.GDS)
d.GDP_imp <- fill_na_with_nearest_avg(d.GDP)
d.Corruption_imp <- fill_na_with_nearest_avg(d.Corruption)
d.RR_imp<-fill_na_with_nearest_avg(d.RR)
d.CC_imp<-fill_na_with_nearest_avg(d.CC)
d.ED_imp <- fill_na_with_nearest_avg(d.ED)

# Save as _UNimp the unimputed sets
d.Adoption_UNimp <- d.Adoption
d.Inflation_UNimp <- d.Inflation
d.GDS_UNimp <- d.GDS
d.GDP_UNimp <- d.GDP
d.Corruption_UNimp <- d.Corruption
d.RR_UNimp <- d.RR
d.CC_UNimp <- d.CC
d.ED_UNimp <- d.ED

```

### Interpretation of Missing Data Pattern Figure

The figure consists of horizontal bars, one for each of the configurations of missing data. Where blue represents a present year and red an absent year. The top headings represent the row headings. The number to the left of each bar represents the number of times a configuration of missing data is represented in the dataset. The number on the right represents the number of missing data points in a single observation, for a particular configuration of missing data. The numbers in the bottom footer, represent the number of times a particular feature is missing across the dataset. The number at the bottom right represents the total number of missing variables for each dataset.

```{r creating-long-func, echo = F}
convert_to_long <- function(datasets) {
  for (dataset_name in datasets) {
    if (exists(dataset_name, envir = .GlobalEnv)) {
      dataset <- get(dataset_name, envir = .GlobalEnv)
      
      # Extract the part of the dataset name after "d." and before "_imp"
      value_name <- gsub("^d\\.|_imp$", "", dataset_name)
      
      long_name <- paste0(dataset_name, "_long")
      
      long_data <- dataset %>% 
        pivot_longer(cols = -Country,  
                     names_to = "Year", 
                     values_to = value_name) %>%  
        mutate(Year = trimws(Year)) %>%  # Remove leading/trailing spaces
        mutate(Year = gsub("[^0-9]", "", Year)) %>%  # Remove non-numeric characters
        mutate(Year = as.integer(Year))  # Convert cleaned Year to integer

      assign(long_name, long_data, envir = .GlobalEnv)
    } else {
      message(paste("Dataset", dataset_name, "does not exist in the environment."))
    }
  }
}
```

```{r, applying-long-func, echo = F}
dataset_imp_list <- c( "d.CC_imp","d.Corruption_imp","d.ED_imp", "d.GDP_imp", "d.GDS_imp", "d.Inflation_imp","d.RR_imp", "d.Adoption_imp")
convert_to_long(dataset_imp_list)
```

```{r left-join, echo = F}
d.panel <- d.Adoption_imp_long
d.panel<-d.panel %>% 
  left_join(d.Inflation_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.GDS_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.GDP_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.Corruption_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.RR_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.CC_imp_long, by= c( "Country","Year")) %>% 
  left_join(d.ED_imp_long, by= c( "Country","Year")) 

# I performed "Stichproben" for these to check the aggregation is in line with the original data.

```

### Yeo - Johnson Transformation

The Yeo - Johnson transformation is done using the package {caret}. First, a function is used to estimate the lambda parameters of transformable predictor variables. Table \@ref(tab:YJ-lambda-tbl) shows an overview of the estimates. It should be noted that the Country Name variable was not transformed as it is not numeric and the Year will be treated as categorical. The data on CC has insufficient data diversity naturally, so this data was altered with perturbations drawn from a distribution with mean 0 and a standard deviation equal to 6% the range of the capital controls variable (the smallest possible value while still allowing the transformation).

[[**need reference, define how it will be treated**]{.underline}].

```{r YJ, echo = F}
set.seed(42)
num_cols <- d.panel[,!names(d.panel) %in% c("Country","Year")] # selecting the numeric columns only
num_cols$CC <- num_cols$CC + rnorm(nrow(num_cols), mean = 0,sd= 0.06) # adding  pertubations

preProcess_model <- preProcess(num_cols, method = "YeoJohnson", center = FALSE, scale = FALSE)
d.panel_YJ <- predict(preProcess_model, num_cols)
d.panel_YJ$Year <-d.panel$Year
d.panel_YJ$Country <-d.panel$Country

#d.panel_YJ$CC <-d.panel$CC

d.panel_YJ <- d.panel_YJ %>% 
  select(Country,Year, everything())#,CC)

```

```{r YJ-lambda-tbl, echo = F}
# Convert lambda values to a dataframe and add variable names
lambda_YJ <- data.frame(
  Variable = colnames(num_cols),  # Get variable names from num_cols
  Lambda = preProcess_model$yj  # Extract lambda values
)

# Round the lambda values to 4 decimal places
lambda_YJ$Lambda <- round(lambda_YJ$Lambda, 4)

# Rename the column
colnames(lambda_YJ) <- c("Variable", "Lambda Estimate")


kable(lambda_YJ, caption = "Yeo-Johsnon Transformations Lambda Estimates (Model 2)", format = "latex", booktabs = TRUE, escape = FALSE, row.names = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs"))
  #column_spec(2, width = "3cm")  # Adjust column width for alignment

```

\newpage

# Exploratory Data Analysis

This section makes exploratory analysis of the data to both understand descriptive statics and to prepare for inferential statistics by checking if assumptions for certain models are met. Please note, this entire section is performed on the imputed data, and in the case of the CC data in the Yeo - Johnson transformed datasets, the data including the perturbations added to allow the transformation of that variable.

## Distribution: Histograms

[[**Explain what histograms are and the benefit they bring**]{.underline}]

```{r hist_fun, echo = F}
plot_side_by_side_histogram <- function(datasets, var_name, labels, bins = 30) {
  # Check if the number of datasets matches the number of labels
  if (length(datasets) != length(labels)) {
    stop("Number of datasets must match number of labels.")
  }
  
  # Compute Shapiro-Wilk normality test p-values
  normality_results <- sapply(datasets, function(df) {
    p_value <- shapiro.test(df[[var_name]])$p.value
    if (p_value < 0.001) {
      return("< 0.001")  # Avoids scientific notation
    } else {
      return(sprintf("%.3f", p_value))  # Round to 3 decimals
    }
  })
  
  # Combine datasets into a long format dataframe
  combined_data <- bind_rows(
    lapply(seq_along(datasets), function(i) {
      datasets[[i]] %>%
        select(all_of(var_name)) %>%
        mutate(Transformation = paste0(labels[i], "\nShapiro p = ", normality_results[i]))
    }),
    .id = "Source"
  )
  
  # Plot the histograms side by side
  ggplot(combined_data, aes_string(x = var_name)) +
    geom_histogram(fill = "blue", alpha = 0.6, bins = bins) +
    facet_wrap(~Transformation, scales = "free") +  # Facet by transformation (with normality p-value)
    labs(x = var_name, y = "Count") +
    theme_minimal() +
    theme(strip.text = element_text(size = 12, face = "bold"))
}

```

**Adoption Data**

Figure \@ref(fig:hist-Adoption) shows the distribution of the Adoption data in both transformations.

\FloatBarrier

```{r hist-Adoption, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.Adoption", warning=F}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "Adoption",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

```{=tex}
\FloatBarrier
\clearpage
```
**Inflation**

Figure \@ref(fig:hist-inf) shows the histograms for the inflation data.

\FloatBarrier

```{r hist-inf, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.Inflation"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "Inflation",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

```{=tex}
\FloatBarrier
\clearpage
```
**Gross Domestic Savings**

Figure \@ref(fig:hist-GDS) shows the histograms for Gross Domestic Savings.

\FloatBarrier

```{r hist-GDS, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.GDS"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "GDS",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

```{=tex}
\FloatBarrier
\clearpage
```
**Gross Domestic Product**

Figure \@ref(fig:hist-GDP), shows the histograms for Gross Domestic Product.

\FloatBarrier

```{r hist-GDP, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.GDP"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "GDP",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)

```

```{=tex}
\FloatBarrier
\clearpage
```
**Sins**

Figure \@ref(fig:hist-Corruption), which shows the distribution of the Sins Proxy, Corruption.

\FloatBarrier

```{r hist-Corruption, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.Corruption"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "Corruption",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

```{=tex}
\FloatBarrier
\clearpage
```
**Remittances**

Figure \@ref(fig:hist-RR) shows the histograms for Remittances.

\FloatBarrier

```{r hist-RR, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.RR"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "RR",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

```{=tex}
\FloatBarrier
\clearpage
```
**Capital Controls**

Figure \@ref(fig:hist-CC) show the histograms for Capital Controls.

\FloatBarrier

```{r hist-CC, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.CC"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "CC",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 5)
```

```{=tex}
\FloatBarrier
\clearpage
```
**External Debt to GDP**

Figure \@ref(fig:hist-ED-GDP) show that the distribution for External Debt to GDP.

\FloatBarrier

```{r hist-ED-GDP, echo = F, message = F, fig.width=6, fig.height=4, fig.cap="Histograms for d.ED"}
plot_side_by_side_histogram(datasets = list(d.panel,d.panel_YJ),var_name = "ED",labels = c("No Transformation", "Yeo - Johnson Transormation"), bin = 30)
```

\FloatBarrier

## Overview Distribution

Table \@ref(tab:overview-skew-no-transformations) shows an overview of the distributions of the un-transformed variables.

```{r overview-skew-no-transformations, echo = F, results = "asis"}
# Create the data frame
distribution_table <- data.frame(
  Variable = c("Adoption", "Inflation", "GDS", "GDP", "CC", "RR", "ED"),
  Distribution = c("long right tail", "long right tail", "quasi-normal",
                   "right skew", "right skew", "right skew", "right skew")
)

# Print the table and ensure it holds position
kable(distribution_table, caption = "Overview of Distributions (No Transformations)") %>%
  kable_styling(latex_options = "HOLD_position")

```

## Correlation Matrix

```{r corr-heatmap-create, echo = F}
d.panel_numeric <- subset(d.panel, select = -c(Country, Adoption))
corr_matrix <- cor(d.panel_numeric)
melted_corr<-melt(corr_matrix)

#ggplot(melted_corr, aes(x=Var1, y=Var2, fill=value)) +
#  geom_tile() +
#  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, 
#                       limits = c(-1, 1)) +  # Set limits to force full color range
#  theme_minimal() +
#  labs(title="Correlation Heatmap", fill="Correlation") +
#  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Create the heatmap with correlation values
corr_heatmap<-ggplot(melted_corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +  # Heatmap tiles
  geom_text(aes(label = round(value, 2)), color = "black", size = 4) +  # Add correlation values
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, 
                       limits = c(-1, 1)) +  # Ensure color scale is -1 to 1
  theme_minimal() +
  labs(fill="Correlation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

melted_corr_wo1 <- melted_corr[melted_corr$value != 1, ]
med_corr <- round(median(abs(melted_corr_wo1$value)),2)
max_corr <- round(max(abs(melted_corr_wo1$value)),2)
```

A correlation matrix displays visually a relationship between two continuous variables. It is important to evaluate this, since an assumption behind many some statistical models, particularily regression is that there is no **multicollinearity** [@statisticssolutions]. Figure \@ref(fig:corr-heatmap-show) shows a correlation heatmap for independent variables, the numbers inside of the cells (as well as their coloring, as indicated on the heatmap) shows the pearson correlation coefficient between two independent variables in the datasets. The variables associates with each correlation can be identified by looking at the row and column of that cell. The plot shows that most correlations are weak to moderate, with a median absolute value of correlations equal to `r med_corr` and a maximum equal to `r max_corr`.[^3] This indicates that multicollinearity will not be a problem since even the largest absolute value is below the 0.8 threshold specified in @statisticssolutions. The diagonal values of 1 represent to correlation of each variable with itself, which is not directly relevant. Please note, each correlation is represented twice (one with a given variable on the x-axis and once on the y-axis.

[^3]: The summary statistics present here were calculated without the correlation of variables with themselves (diagonal 1s in Figure \@ref(fig:corr-heatmap-show))

\FloatBarrier

```{r corr-heatmap-show, echo = F, fig.cap="Correlation Heatmap of Independent Variables", fig.pos="h"}
corr_heatmap
```

\FloatBarrier

## Hausman Test

```{r, echo = F, warning=F, message = F}
library(plm)
m3.fixed <- plm(Adoption~Inflation+GDS+GDP+Corruption+RR+CC+ED, data = d.panel, index=c("Country","Year"), model = "within")
m3.random <- plm(Adoption~Inflation+GDS+GDP+Corruption+RR+CC+ED, data = d.panel, index=c("Country","Year"), model = "random")

```

```{r ht-tbl, echo = F, warning = F, message = F }
ht <- phtest(m3.fixed,m3.random)
ht_df <- data.frame(
  chisq = as.numeric(ht$statistic),
  p_Value = ht$p.value
  #Formula = ht$data.name
)

kable(ht_df, format = "latex", booktabs = TRUE, caption = "Results of Hausman Test Fixed vs. Random Effects") %>%
  kable_styling(latex_options = c("hold_position"))
```

To determine whether a fixed or random effects model will be used, the Hausman Test can be used, provided that the observations are randomly drawn from the population [@dougherty2011; @qin2023]. Unfortunately, the authors of the @statista_adoption dataset do not disclose the methodology for selecting the countries. However, since the survey was part of the **Global** Consumer Survey, it can be assumed that the authors made an effort to include a diversity of countries from an economic and political standpoint [@statista2020]. Certainly while looking at the map of countries included in the survey (Figure \@ref(fig:fig-worldmap-selected)), no immediate category of country is apparent on factors like geographical region, political system of development level.

Table \@ref(tab:ht-tbl) shows the result of the Hausman test, indicating that the $H_0$ should be rejected at the low p-value of `r ht$p.value` and that a fixed effect model is preferred [@dougherty2011; @qin2023].

\newpage

# Results

```{r, echo = F}
res2var <- function(df) {
  for (i in 1:nrow(df)) {
    coefficient <- gsub("[()]", "", df$coefficient[i])  # Remove special characters like parentheses
    coefficient <- gsub(" ", "_", coefficient)  # Replace spaces with underscores
    
    assign(paste0("s.1.", coefficient, ".estimate"), df$estimate[i], envir = .GlobalEnv)
    assign(paste0("s.1.", coefficient, ".std.error"), df$std.error[i], envir = .GlobalEnv)
    assign(paste0("s.1.", coefficient, "t.statistic"), df$t.statistic[i], envir = .GlobalEnv)
    assign(paste0("s.1.", coefficient, ".p.value"), df$p.value[i], envir = .GlobalEnv)
}}
```

This section discussed the statistical results of the models. The overviews are presented as Tables and Figures. The definition of the metrics is discussed, however only key relevant results discussed.

## Model 1: Linear Regression No Transformation

```{r, echo = F}
m.1 <- lm(Adoption~as.factor(Year)+Inflation+GDS+GDP+Corruption+RR+CC+ED, data = d.panel)
```

```{r model1-presentation-coefficients, echo = F}
library(knitr) #gotta see if runs without loading here
library(kableExtra)
library(broom)
library(dplyr)

# Format the model summary
s.1 <- tidy(m.1) %>%
  mutate(
    term = gsub("_", "\\\\_", term),  # Escape underscores for LaTeX
    estimate = sprintf("%.2f", estimate),  # 2 decimal places
    std.error = sprintf("%.2f", std.error),  # 2 decimal places
    statistic = sprintf("%.2f", statistic),  # 2 decimal places
    p.value = sprintf("%.4f", p.value)  # 4 decimal places
  ) %>% 
  rename(coefficient = term) %>% 
  rename(t.statistic = statistic)

# Create LaTeX table
kable(s.1, caption = "Model 1 Coefficients", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs")) %>%
  column_spec(2, width = "3cm")  # Adjust column width for alignment

res2var(s.1)

```

```{r model1-presentation-summary, echo = F}
summary_stats <- summary(m.1)

# Extract residuals
residuals_summary <- summary(m.1$residuals)

# Create a dataframe for model fit statistics
model_summary <- data.frame(
  Metric = c("Residual Standard Error", 
             "Multiple R-squared", 
             "Adjusted R-squared", 
             "F-statistic", 
             "Degrees of Freedom (Model)", 
             "Degrees of Freedom (Residuals)",
             "P.value",
             "Residual Min",
             "Residual Q1",
             "Residual Median",
             "Residual Q3",
             "Residual Max"),
  Value = c(sprintf("%.2f", summary_stats$sigma),
            sprintf("%.4f", summary_stats$r.squared),
            sprintf("%.4f", summary_stats$adj.r.squared),
            sprintf("%.2f", summary_stats$fstatistic[1]),
            summary_stats$df[1],
            summary_stats$df[2],
            sprintf("%.4f", pf(summary_stats$fstatistic[1], 
                               summary_stats$fstatistic[2], 
                               summary_stats$fstatistic[3], 
                               lower.tail = FALSE)),
            sprintf("%.2f", residuals_summary[1]),  # Min
            sprintf("%.2f", residuals_summary[2]),  # Q1
            sprintf("%.2f", residuals_summary[3]),  # Median
            sprintf("%.2f", residuals_summary[5]),  # Q3
            sprintf("%.2f", residuals_summary[6]))  # Max
)

# Create a formatted LaTeX table using kable
kable(model_summary, caption = "Model 1 Fit Summary", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs"))
```

```{r model1predres, echo = F, fig.cap="Model 1: Scatterplot Showing Predicted vs. Residuals", results = "asis", fig.width=5, fig.height=4}

# Create a data frame with predictions and residuals
df <- data.frame(
  Predicted = predict(m.1),
  Residuals = residuals(m.1)
)

# Generate the ggplot
ggplot(df, aes(x = Predicted, y = Residuals)) +
  geom_point() +  # Scatter plot
  geom_hline(yintercept = 0, linetype = "dashed", color = "lightblue", size = 1) +  # Horizontal line at 0
  ylim(-30,30)+
  labs(x = "Predicted Values", y = "Residuals") +
  theme_minimal()
```

```{r model1residualhistograms, fig.cap= "Model 1: Histogram of Predicted Values", fig.width=5, fig.height=4, echo =F, fig.pos="H"}
m.1_residuals <- data.frame(residuals = m.1$residuals)
# Create histogram of residuals
ggplot(m.1_residuals, aes(x = residuals)) +
  geom_histogram(bins = 50, fill = "blue") +
  labs(title = "", x = "Residuals", y = "Frequency") +
  theme_minimal()

```

Linear Regression models are one of the simplest inferential statistical tools out there. Nevertheless they can provide detailed information on the relationships between variables. As discussed already, this model considers the years as categorical variables while the other dependent variables are modelled as continuous.

Table \@ref(tab:model1-presentation-coefficients) shows the results of this model in terms of the coefficients of the independent variables. The estimates for categorical values (as.factor) represent linear offsets in the **estimate** for the dependent variable for the different years, relative to the reference year, holding other predictors constant. While the estimates for continuous values represent the expected change in the dependent variable as a result of a one unit change in the respective variable, again while holding other variables constant. The **std.error** is an approximation of the deviation of the coefficient from its estimate. Combined with the estimate, it can be useful to get confidence intervals for coefficients, and more importantly check that the direction of effect does not make the estimate change direction within a certain confidence interval. For example if the 95% confidence interval for an estimate is $2\pm (1.96*3)$ then we cannot even say which direction in response variable happens are a result of a one unit change in independent variable. The **t.statistic** quantifies this idea by calculating $\frac{estimate}{std.error}$ giving a value where a higher number represents greater confidence in the estimate and vice-versa. The **P.value** is the result of a hypothesis test, it gives an indication of how many times values as extreme as the one seen in the data would be seen if the null hypothesis was true (no relationship between predictor and dependent variable). A value of 0.1 means 1 in 10 times values as extreme as the one in the data would be seen if the null hypothesis is true. It therefore measures confidence in the estimates and is often used as a deciding factor whether to conclude a predictor variable effects the dependent variable [@barbanti2024; @thieme2021; @w3schools].

Table \@ref(tab:model1-presentation-summary) shows the model's summary statistics. The residuals are the difference between the actual values and the predicted values. **Residual Standard Error** represents the mean deviation of the true value from the value predicted by the model. The value for **Multiple R-Squared** represent the proportion of variation in dependent explained by the model's independent values. **Adjusted R-Squared** is similar, but it adjusts the proportion explained for the amount of independent variables used. The **F-Statistic** and **p.value** do not provide an interesting additional input in the context of a multivariate regression, when we are anyways looking at the coefficients of the variables, as it is used to determine whether any of the predictors have an impact on the dependent variables [@greenwood2022; @thieme2021]. The Residual Degrees of Freedom represents the the difference between the number of regression coefficients and the number of observations. The **Degrees of Freedom (Model)** and **(Residual)** represent the number of parameters estimated and $n_{observations}-df_{model}$ respectively. The summary statistics **Residual Min**, **Residual Quartile 1**, **Residual Median**, **Residual Quartile 3** and **Residual Max** describe summary statistics of all the residuals for individual observations across years. The numbers indicate that the error is evenly distributed between over and under estimation of the cryptocurrency adoption [@thieme2021].

Figure \@ref(fig:model1predres) shows the scatterplot of the values predicted by the model versus the residuals. The deviation of predictions on the higher end of the range towards higher residuals indicates that the model is not equally good at predicting throughout at ranges of the data. This is known as **heteroskedasticity**.

Figure \@ref(fig:model1residualhistograms) shows a distribution of the residual term in the first model.

[[**Small Model Summary!**]{.underline}]

## Model 2: Linear Regression Transformations

```{r, echo = F}
m.2 <- lm(Adoption~as.factor(Year)+Inflation+GDS+GDP+Corruption+RR+CC+ED, data = d.panel_YJ)
```

```{r model2-presentation-coefficients, echo = F}

# Format the model summary
s.2 <- tidy(m.2) %>%
  mutate(
    term = gsub("_", "\\\\_", term),  # Escape underscores for LaTeX
    estimate = sprintf("%.2f", estimate),  # 2 decimal places
    std.error = sprintf("%.2f", std.error),  # 2 decimal places
    statistic = sprintf("%.2f", statistic),  # 2 decimal places
    p.value = sprintf("%.4f", p.value)  # 4 decimal places
  ) %>% 
  rename(coefficient = term) %>% 
  rename(t.statistic = statistic)

# Create LaTeX table
kable(s.2, caption = "Model 2 Coefficients", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs")) %>%
  column_spec(2, width = "3cm")  # Adjust column width for alignment

res2var(s.2)
```

Table \@ref(tab:model2-presentation-coefficients) shows the resulting coefficients of the second model.

```{r model2-presentation-summary, echo = F}
summary_stats <- summary(m.2)

# Extract residuals
residuals_summary <- summary(m.2$residuals)

# Create a dataframe for model fit statistics
model_summary <- data.frame(
  Metric = c("Residual Standard Error", 
             "Multiple R-squared", 
             "Adjusted R-squared", 
             "F-statistic", 
             "Degrees of Freedom (Model)", 
             "Degrees of Freedom (Residuals)",
             "P.value",
             "Residual Min",
             "Residual Q1",
             "Residual Median",
             "Residual Q3",
             "Residual Max"),
  Value = c(sprintf("%.2f", summary_stats$sigma),
            sprintf("%.4f", summary_stats$r.squared),
            sprintf("%.4f", summary_stats$adj.r.squared),
            sprintf("%.2f", summary_stats$fstatistic[1]),
            summary_stats$df[1],
            summary_stats$df[2],
            sprintf("%.4f", pf(summary_stats$fstatistic[1], 
                               summary_stats$fstatistic[2], 
                               summary_stats$fstatistic[3], 
                               lower.tail = FALSE)),
            sprintf("%.2f", residuals_summary[1]),  # Min
            sprintf("%.2f", residuals_summary[2]),  # Q1
            sprintf("%.2f", residuals_summary[3]),  # Median
            sprintf("%.2f", residuals_summary[5]),  # Q3
            sprintf("%.2f", residuals_summary[6]))  # Max
)

# Create a formatted LaTeX table using kable
kable(model_summary, caption = "Model 2 Fit Summary", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs"))
```

Table \@ref(tab:model2-presentation-summary) shows the summary statistics of the Linear Regression with Yeo - Johnson transformation applied.

\FloatBarrier

```{r model2predres, echo = F, fig.cap="Model 2: Scatterplot Showing Predicted vs. Residuals", results = "asis", fig.width=5, fig.height=4, fig.pos="H"}

# Create a data frame with predictions and residuals
df <- data.frame(
  Predicted = predict(m.2),
  Residuals = residuals(m.2)
)

# Generate the ggplot
ggplot(df, aes(x = Predicted, y = Residuals)) +
  geom_point() +  # Scatter plot
  geom_hline(yintercept = 0, linetype = "dashed", color = "lightblue", size = 1) +  # Horizontal line at 0
  ylim(-1,1)+
  labs(x = "Predicted Values", y = "Residuals") +
  theme_minimal()
```

Figure \@ref(fig:model2predres) shows the values predicted by the model vs. the actual values. It is on the **transformed scale**. [[**DESCRIBE**]{.underline}]

```{r model2residualhistograms, fig.cap= "Model 2: Histogram of Predicted Values", fig.width=5, fig.height=4, echo =F, fig.pos="H"}
m.2_residuals <- data.frame(residuals = m.2$residuals)
# Create histogram of residuals
ggplot(m.2_residuals, aes(x = residuals)) +
  geom_histogram(bins = 50, fill = "blue") +
  labs(title = "", x = "Residuals", y = "Frequency") +
  theme_minimal()
```

Figure \@ref(fig:model2residualhistograms) shows the histogram of the residuals in the model. [[**DESCRIBE**]{.underline}]

\FloatBarrier

## Model 3: Fixed Effects

Table \@ref(tab:model3-presentation-coefficients) shows the coefficients of Model 3.

```{r model3-presentation-coefficients, echo = F}

# Format the model summary
s.3 <- tidy(m3.fixed) %>%
  mutate(
    term = gsub("_", "\\\\_", term),  # Escape underscores for LaTeX
    estimate = sprintf("%.2f", estimate),  # 2 decimal places
    std.error = sprintf("%.2f", std.error),  # 2 decimal places
    statistic = sprintf("%.2f", statistic),  # 2 decimal places
    p.value = sprintf("%.4f", p.value)  # 4 decimal places
  ) %>% 
  rename(coefficient = term) %>% 
  rename(t.statistic = statistic)

# Create LaTeX table
kable(s.3, caption = "Model 3 Coefficients", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs")) %>%
  column_spec(2, width = "3cm")  # Adjust column width for alignment
```

Table \@ref(tab:model3-presentation-summary) shows the fit of Model 3.

```{r model3-presentation-summary, echo = F}
summary_stats <- summary(m3.fixed)

# Extract residuals
residuals_summary <- summary(m3.fixed$residuals)

# Create a dataframe for model fit statistics
model_summary <- data.frame(
  Metric = c(#"Residual Standard Error", 
             "Multiple R-squared", 
             "Adjusted R-squared", 
             "F-statistic", 
             "Degrees of Freedom (Model)", 
             "Degrees of Freedom (Residuals)",
             "P.value",
             "Residual Sum of Squares",
             "Between Entities",
             "Between Time"),

  Value = c(#sprintf("%.2f", summary_stats$sigma), #1
            sprintf("%.4f", summary_stats$r.squared[1]), #2
            sprintf("%.4f", summary_stats$r.squared[2]), #3
            sprintf("%.2f", summary_stats$fstatistic$statistic), #4
            summary_stats$df[1], #5
            summary_stats$df[2], #6
            sprintf("%.4f", summary_stats$fstatistic$p.value), #7
            sprintf("%.2f", residuals_summary[1]),  # Sum of Sq #8
            sprintf("%.2f", residuals_summary[2]),  # In between id #9
            sprintf("%.2f", residuals_summary[3])))

# Create a formatted LaTeX table using kable
kable(model_summary, caption = "Model 3 Fit Summary", format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(latex_options = c("hold_position", "booktabs"))
```

Figure \@ref(fig:model3predres) shows the predicted versus residuals of this model.

```{r model3predres, echo = F, fig.cap="Model 3: Scatterplot Showing Predicted vs. Residuals", results = "asis", fig.width=5, fig.height=4, fig.pos="H"}

# Create a data frame with predictions and residuals
df <- data.frame(
  Predicted = as.numeric(predict(m3.fixed)),
  Residuals = as.numeric(residuals(m3.fixed))
)

# Generate the ggplot
ggplot(df, aes(x = Predicted, y = Residuals)) +
  geom_point() +  # Scatter plot
  geom_hline(yintercept = 0, linetype = "dashed", color = "lightblue", size = 1) +  # Horizontal line at 0
  ylim(-20,20)+
  labs(x = "Predicted Values", y = "Residuals") +
  theme_minimal()
```

Figure \@ref(fig:model3residualhistograms) shows the distribution of the residuals.

```{r model3residualhistograms, fig.cap= "Model 3: Histogram of Predicted Values", fig.width=5, fig.height=4, echo =F, fig.pos="H"}
m.3_residuals <- data.frame(residuals = m3.fixed$residuals)
# Create histogram of residuals
ggplot(m.3_residuals, aes(x = residuals)) +
  geom_histogram(bins = 50, fill = "blue") +
  labs(title = "", x = "Residuals", y = "Frequency") +
  theme_minimal()
```

\newpage

# Discussion

## Summary Key Findings

## Comparison With Existing Literature

## Strengths of Approach

## Weaknesses of Approach

Point 1: Despite the inclusion of a variable considering Capital Controls by creating an index, this variable brought considerable challenges from a technical side, as despite theoretically being continuous, with the inclusion of enough variables.

Point 2: Causality not established

## Future Research

Point 1: Lack of data presents an issue in this research field, but the availability is improving when compared to previous approaches (either highly technical (see preliminary study) or far less complete (previous Statista set). Even so, the extreme outliers on the economic end (think Argentina, Venezuela, Zimbabwe) do not produce reliable economic information either.

## Practical and Policy Implications

# Conclusion

\newpage

# Bibliography (APA 7\textsuperscript{th})

::: {#refs}
:::

\newpage

# Appendix 1: Structured List of Literature

This Appendix gives a visual overview of the two main important fields of academic literature for this paper.

## Literature Area 1: Drivers of Currency Substitution

Table \@ref(tab:litreviewCS) below shows the overview of the key literature on currency substitution and in relation to this research. It is a visual representation of the text in section \@ref(currency-substitution-1).

\FloatBarrier

```{r tab:litreviewCS, echo=FALSE, results='asis', fig.pos="H"}
cat("
\\begin{table}[ht]
\\centering
\\includegraphics[width=0.9\\linewidth]{review_CS.png}
\\caption{Visual Summary Currency Substitution Literature}
\\label{tab:litreviewCS}
\\end{table}
")
```

\FloatBarrier

\newpage

## Literature Area 2: Drivers of Cryptocurrency Adoption

Table \@ref(tab:litreviewCA) below shows the overview of the key literature on the adoption of cryptocurrencies in relation to this research. It is a graphical representation of the text in section \@ref(adoption-of-cryptocurrency).

\FloatBarrier

```{r tab:litreviewCA, echo=FALSE, results='asis', fig.pos="H"}
cat("
\\begin{table}[ht]
\\centering
\\includegraphics[width=0.9\\linewidth]{review_bbtc.png}
\\caption{Visual Summary Cryptocurrency Adoption Literature}
\\label{tab:litreviewCA}
\\end{table}
")
```

\FloatBarrier

\newpage

# Appendix 2: Missing Data Patterns

This section shows the Missing patterns for the pre-imputation data of all variables with the exception of *d.CC*, as this was already presented in the section \@ref(missing-data-structure). Please see \@ref(interpretation-of-missing-data-pattern-figure) for a guide on how to interpret the figure.

\FloatBarrier

```{r md-Adoption, echo = F, fig.cap="Missing Pattern d.Adoption", message= F,  results="hide", fig.width=5, fig.height=4}
md.pattern(d.Adoption_UNimp[,-1])
```

```{r md-GDS, echo = F, fig.cap="Missing Pattern d.GDS", fig.width=5, fig.height=4, message= F,  results="hide"}
md.pattern(d.GDS_UNimp[,-1])
```

```{r md-GDP, echo = F, fig.cap="Missing Pattern d.GDP", fig.width=5, fig.height=4, message= F,  results="hide"}
md.pattern(d.GDP_UNimp[,-1])
```

```{r md-Corruption, echo = F, fig.width=5, fig.height=4, fig.cap="Missing Pattern d.Corruption", message= F,  results="hide"}
md.pattern(d.Corruption_UNimp[,-1])
```

```{r md-RR, echo = F, fig.cap="Missing Pattern d.RR", fig.width=5, fig.height=4, message= F,  results="hide"}
md.pattern(d.RR_UNimp[,-1])
```

```{r md-ED-GDP, echo=F, fig.cap="Missing Pattern d.ED", fig.width=5, fig.height=4, message = F, results = "hide"}
md.pattern(d.ED_imp[,-1])
```

\FloatBarrier

\clearpage

\newpage

# Appendix 3: AI Disclosure

I made extensive use of AI, particularly ChatGPT for this project. The main, but not exhaustive uses were the following:

-   Spelling / Reviewing and writing drafts of sections where I needed inspiration to get started.

-   R Studio, Markdown and LaTex issues and code suggestions.

-   Scanning Long Documents for parts that I needed.

-   Methodology "Research", although I verified everything using proper academic sources.

-   Cover Page: Used Canva's AI Image Generation to create the cover photo.

\newpage

# Appendix 4: Declaration of Originality

I hereby confirm that I Alec Vayloyan (born 02.06.2000, Bern),  have written this Thesis independently and without the help of any third party,  have provided all the sources and cited all the literature I used,  will protect the confidentiality of the client and respect the copyright regulations of Lucerne University of Applied Sciences and Arts.

\newpage

# Appendix 5: GitHub Access

This project and associated data can be found on the following public GitHub Repository: [[**INSERT GITHUB REPOSITORY LINK**]{.underline}]

```{r, eval = F, echo = F}


plot_summary_boxplots <- function(data, title = NULL) {
  library(ggplot2)
  library(tidyr)
  library(dplyr)
  
  # Convert data to long format
  data_long <- pivot_longer(data, cols = everything(), names_to = "Year", values_to = "Value")
  
  # Convert Year to factor for ordering
  data_long$Year <- factor(data_long$Year, levels = unique(data_long$Year))

  # Calculate summary statistics
  summary_stats <- data_long %>%
    group_by(Year) %>%
    summarise(
      Median = median(Value, na.rm = TRUE),
      Mean = mean(Value, na.rm = TRUE),
      Q1 = quantile(Value, 0.25, na.rm = TRUE),
      Q3 = quantile(Value, 0.75, na.rm = TRUE),
      IQR = Q3 - Q1
    )

  # Identify outliers using 1.5 * IQR rule
  outliers <- data_long %>%
    left_join(summary_stats, by = "Year") %>%
    filter(Value < Q1 - 1.5 * IQR | Value > Q3 + 1.5 * IQR)

  # NA counts
  na_counts <- data_long %>%
    group_by(Year) %>%
    summarise(NAs = sum(is.na(Value)))

  # Define legend placement
  legend_x <- max(as.numeric(data_long$Year)) + 1.5  # Shift text further right
  symbol_x <- max(as.numeric(data_long$Year)) + 0.8  # Symbol placement
  legend_y_top <- max(data_long$Value, na.rm = TRUE) - 5  # Top position

  # Create the plot
  p <- ggplot(data_long, aes(x = Year, y = Value)) +
    # One-sided violin (left clipped)
    geom_violin(fill = "#CBD5E0", alpha = 0.7, trim = TRUE, draw_quantiles = NULL) +  
    
    # Boxplot (without outliers)
    geom_boxplot(width = 0.2, outlier.shape = NA, fill = "#4A5568", color = "#2D3748", alpha = 0.6) +
    
    # Median as black line
    geom_segment(data = summary_stats, aes(x = Year, xend = Year, y = Median, yend = Median), color = "black", size = 1.5) +
    
    # Mean as dark blue dot
    geom_point(data = summary_stats, aes(x = Year, y = Mean), color = "#2C5282", size = 3, shape = 18) + 
    
    # Outliers as grey dots
    geom_point(data = outliers, aes(x = Year, y = Value), color = "#718096", size = 2, shape = 1) +

    # NA Count Labels
    geom_text(data = na_counts, aes(x = Year, y = max(data_long$Value, na.rm = TRUE) + 5, label = paste0("NA: ", NAs)), 
              color = "black", size = 4) +
    
    # Custom Legend with Properly Aligned Text
    annotate("text", x = legend_x - 0.3, y = legend_y_top, label = "Legend", fontface = "bold", size = 3.5, hjust = 0) +
    
    # Mean
    annotate("point", x = symbol_x, y = legend_y_top - 4, color = "#2C5282", size = 3, shape = 18) +
    annotate("text", x = legend_x, y = legend_y_top - 4, label = "Mean (Blue Diamond)", hjust = 0, size = 4) +
    
    # Median
    annotate("segment", x = symbol_x - 0.2, xend = symbol_x + 0.2, y = legend_y_top - 8, yend = legend_y_top - 8, color = "black", size = 1.5) +
    annotate("text", x = legend_x, y = legend_y_top - 8, label = "Median (Black Line)", hjust = 0, size = 4) +
    
    # IQR Boxplot
    annotate("rect", xmin = symbol_x - 0.3, xmax = symbol_x + 0.3, ymin = legend_y_top - 12, ymax = legend_y_top - 16, fill = "#4A5568", alpha = 0.6) +
    annotate("text", x = legend_x, y = legend_y_top - 14, label = "IQR (Grey Box)", hjust = 0, size = 4) +
    
    # Outliers
    annotate("point", x = symbol_x, y = legend_y_top - 18, color = "#718096", size = 2, shape = 1) +
    annotate("text", x = legend_x, y = legend_y_top - 18, label = "Outliers (Grey Circles)", hjust = 0, size = 4) +

    theme_minimal() +
    labs(x = "Year", y = "Value", title = title) +
    theme(
      legend.position = "none",  # Remove default legend
      panel.grid.major.x = element_blank(),  # Remove vertical gridlines
      panel.grid.minor.x = element_blank(),
      plot.margin = margin(10, 150, 10, 10)  # <-- Increase right margin to prevent text cutoff
    )

  print(p)
}

```

```{r, boxcox, eval = F, echo = F}
library(MASS)

# Run Box-Cox transformation and store results
boxcox_result <- boxcox(lm(d.panel$GDS ~ 1), lambda = seq(-2, 2, 0.1), plotit = TRUE)

# Extract lambda values (x-axis) and log-likelihood values (y-axis)
lambda_vals <- boxcox_result$x
log_lik_vals <- boxcox_result$y

# Find the MLE of lambda (the one maximizing log-likelihood)
lambda_mle <- lambda_vals[which.max(log_lik_vals)]

# Compute the 95% confidence interval threshold (log-likelihood max - 0.5)
ci_threshold <- max(log_lik_vals) - 0.5

# Identify lambda values that fall within the confidence interval
lambda_ci <- lambda_vals[log_lik_vals >= ci_threshold]

# Compute the middle value of the confidence interval
lambda_mid <- mean(range(lambda_ci))

# Print results
cat("MLE Lambda:", lambda_mle, "\n")
cat("95% Confidence Interval:", range(lambda_ci), "\n")
cat("Middle of Confidence Interval:", lambda_mid, "\n")


```

```{r, echo = F, eval = F}



# Apply Yeo-Johnson transformation without centering or scaling
preProcess_model <- preProcess(d.panel, method = "YeoJohnson", center = FALSE, scale = FALSE)

# View the transformation details
print(preProcess_model)

# Apply the transformation
transformed_data <- predict(preProcess_model, d.panel)

# Compare original vs transformed
#par(mfrow = c(1,2))
#hist(d.panel$Adoption, main = "Original Data", col = "lightblue", breaks = 20)
#hist(transformed_data$Adoption, main = "Transformed Data", col = "lightgreen", breaks = 20)

```

## Fixed vs. Panel Paste Out:

The data obtained allows for panel data models to be conducted.

-   control for individual heterogenity (unobserved differences between entities) [@qin2023]

[**Fixed Effects Regression**]{.underline}

**Within-groups fixed effects**

-   subtract the mean values of observations values for an individual, removing the unobserved effect. However this has the downside of removing any variable which remains constant for the individuals, removing any explanatory power of that variable [@dougherty2011].

-   The remaining model will loose a degree of freedom for each entity [@dougherty2011]. Meaning that the risk of over fitting is increased.

**First Difference**

-   time series differencing, removes the individual effects but retain the same probelms (less df, removes variables remaining constant. Gives rise to autocorrelation.

**Least Squares Dummy**

-   unobserved effect is being treated as a dummy variable. Is mathematically equivalent to the within group fixed effect.

**Random Effects**

-   Samples should be drawn form the population randomly to ensure unobserved variables can be treated as randomly distributed.

-   Unobserved variables should not be correlated with any of the predictor variables.
